---
title: "Figure Creation"
author: "Emilee Lance"
date: "2024-08-21"
output: pdf_document
---

```{r setup, warning = FALSE, error = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE)
library(seriation)
library(dplyr)
library(purrr)
library(ggplot2)
library(phyloseq)
library(microViz)
library(shiny)
library(multcomp)
library(svglite)
library(tidyverse)
library(ggtext)
library(ggpubr)
library(Maaslin2)
```

# Load the Phyloseq object

The phyloseq object was created by following the DADA2 workflow. Documentation can be found here: ./Code/Analysis/Zebrafish DADA2.Rmd. Phyloseq object saved here: ./Data/ps_object.rds. For this workflow we will be looking only at the Control and Exposed samples for all 4 strains of zebrafish (5D, AB, Casper & TU).

```{r phyloseq object, tidy = TRUE}
ps <- readRDS("C:/Users/emile/OneDrive - Oregon State University/Sharpton Lab/Interstrain_Zebrafish_Gut_Microbiome/Data/ps_object.rds")

ps_exposed <- ps_filter(ps, Treatment == "Exposed")

```
# File Set-up
```{r}
path = "C:/Users/emile/OneDrive - Oregon State University/Sharpton Lab/Interstrain_Zebrafish_Gut_Microbiome/Figures"

ps@sam_data$FillGroup <- ifelse(
  ps@sam_data$Treatment == "Control",
  "Control",
  as.character(ps@sam_data$Strain)
)

ps@sam_data$FillGroup2 <- with(
  ps@sam_data,
  ifelse(
    Infection.Status == "positive",
    as.character(Strain),   # always filled
    ifelse(
      Infection.Status == "negative" & FillGroup != "Control",
      as.character(Strain), # exposed-negative filled
      "Control"             # control-negative unfilled
    )
  )
)

```
#Define Palettes and Theme
```{r}
strain_colors <- c(
  "5D" = "#E63073", 
  "AB" = "#82B7E6", 
  "Casper" = "#FFC107", 
  "TU" = "#44B5A2"
)

fill_colors <- c(
  "Control" = "white", 
  "5D" = "#E63073", 
  "AB" = "#82B7E6", 
  "Casper" = "#FFC107", 
  "TU" = "#44B5A2"
)
shape_values <- c("Control" = 21, "Exposed" = 21)

custom_theme <- theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95")
  )
```

#Physiology
```{r}
# Weight by strain and pathology results with conditional fill per treatment
# Not included in the paper
strain_pathology_weight_boxplot <- ps@sam_data %>%
  ggplot(aes(y = Weight_mg, x = Strain)) +

  # Boxplot with strain-colored outlines
  geom_boxplot(
    width = 0.7,
    aes(color = Strain),
    outlier.shape = NA
  ) +

  # Jittered points: outline = strain, fill = FillGroup, shape = Treatment
  geom_jitter(
    aes(
      color = Strain,
      fill  = FillGroup,
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75,
    show.legend = TRUE
  ) +

  # Named fill scale so colors map correctly
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +

  # Shapes for exposure group
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Control" = 21, "Exposed" = 21)
  ) +

  # Strain color outlines
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +

  # Legend cleanup
  guides(
    fill = "none",  # hide fill legend if desired
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = c("white", "black"),  # Control = unfilled, Exposed = filled
        color = "black"
      )
    )
  ) +

  theme_bw() +
  facet_wrap(vars(Infection.Status)) +
  theme(custom_theme) +
  labs(y = "Weight (mg)")
strain_pathology_weight_boxplot
ggsave(file = "strain_pathology_weight_boxplot.svg", plot = strain_pathology_weight_boxplot,width = 3.5, height =3, path=path)

# Weight by strain and pathology results with conditional fill per treatment
# Format significance indicators

weight_by_path_df <- pairs_df %>%
  separate(contrast, into = c("group1", "group2"), sep = " - ") %>%
  mutate(
    p.label = case_when(
      p.value <= 0.001 ~ "***",
      p.value <= 0.01  ~ "**",
      p.value <= 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )


weight_by_path_df <- weight_by_path_df %>%
  group_by(Strain) %>%
  mutate(
    y.position = max(sam_data$Weight_mg, na.rm = TRUE) +
                 seq(1, 2, length.out = n())  # only a couple comparisons
  ) %>%
  ungroup()

# Figure 2C
strain_pathology_weight_boxplot_2 <- ps@sam_data %>%
  as.data.frame() %>%
  ggplot(aes(y = Weight_mg, x = Infection.Status)) +

  # Boxplot outlines by strain
  geom_boxplot(
    width = 0.7,
    aes(color = Strain),
    outlier.shape = NA
  ) +

  # Points: outline = strain, fill = FillGroup2, shape = Treatment
  geom_jitter(
    aes(
      color = Strain,
      fill  = FillGroup2,
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75,
    show.legend = TRUE
  ) +

  # Correct, named fill scale
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +

  # Shapes for exposure group
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Control" = 21, "Exposed" = 21)
  ) +

  # Strain outline colors
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +

  # Legend cleanup
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = c("white", "black"),
        color = "black"
      )
    )
  ) +

  theme_bw() +
  facet_wrap(vars(Strain), ncol = 4) +
  theme(custom_theme) +
  labs(y = "Weight (mg)", x = "Infection Status") +
  stat_pvalue_manual(
    weight_by_path_df,
    label = "p.label",
    xmin = "group1",
    xmax = "group2",
    y.position = "y.position",
    tip.length = 0.01,
    size = 2.5
  )


strain_pathology_weight_boxplot_2
ggsave(file = "strain_pathology_weight_boxplot_2.svg", plot = strain_pathology_weight_boxplot_2,width = 7, height =4, path=path)

#Worm burden of exposed fish by strain
# Figure 2A
worm_burden_by_strain<- ps_exposed@sam_data %>%
     ggplot(aes(y = Total.Worm.Count, x = Strain)) +
     geom_boxplot(width = 0.7, aes(color = Strain), outlier.shape = NA) +
     geom_jitter(aes(color = Strain), size = 1, stroke = 0.75, show.legend = T) +
        scale_color_manual(
       name = "Strain",
       values = c(
          "5D" = "#E63073", 
          "AB" = "#82B7E6", 
          "Casper" = "#FFC107", 
          "TU" = "#44B5A2"
        )
      ) +
     theme_bw() +
     labs(y = "Total Worm Count") +
    theme(custom_theme) +
  stat_pvalue_manual(burden_strain_pval_df, 
                     label = "p.label",
                             y.position = c(20,17, 22.5,19, 21, 16),
                             size = 1.5,
                             label.size = 2.5,
                             bracket.size = 0.5,
                             hide.ns = F)
worm_burden_by_strain
ggsave(file = "worm_burden_by_strain.svg", plot = worm_burden_by_strain,width = 3.5, height =3, path=path)


# Worm burden of exposed fish by DPE
# Figure 2B
worm_burden_by_dpe<- ps_exposed@sam_data %>%
     ggplot(aes(y = Total.Worm.Count, x = DPE)) +
     geom_boxplot(width = 0.7,  outlier.shape = NA) +
     geom_jitter(aes(color = Strain), size = 1, stroke = 0.75, show.legend = T) +
        scale_color_manual(
       name = "Strain",
       values = c(
          "5D" = "#E63073", 
          "AB" = "#82B7E6", 
          "Casper" = "#FFC107", 
          "TU" = "#44B5A2"
        )
      ) +
     theme_bw() +
     labs(y = "Total Worm Count") +
    theme(custom_theme) +
  stat_pvalue_manual(burden_anova_dpe_df, 
                     label = "p.label",
                             y.position = 17,
                             size = 1.5,
                             label.size = 2.5,
                             bracket.size = 0.5,
                             hide.ns = F)
worm_burden_by_dpe
ggsave(file = "worm_burden_by_dpe.svg", plot = worm_burden_by_dpe,width = 3.5, height =3, path=path)

```

#Alpha Diversity
```{r}
#facet wrapped boxplot of both alpha diversity metrics without treatment
ps_long <- ps %>%
  ps_calc_diversity(rank = "Genus", index = "shannon") %>%
  ps_calc_richness(rank = "Genus", index = "observed") %>%
  samdat_tbl() %>%
  pivot_longer(cols = c("shannon_Genus", "observed_Genus"), names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = recode(Metric, 
                         "shannon_Genus" = "Shannon Entropy", 
                         "observed_Genus" = "# Observed Genera"))



#Boxplot showing alpha diversity metrics for all samples with fill by exposure group
# Figure S1A
strain_alpha_boxplot <- ps_long %>%
  ggplot(aes(y = Value, x = Strain)) +
  
  geom_boxplot(
    width = 0.7,
    aes(color = Strain),
    outlier.shape = NA
  ) +
  
  geom_jitter(
    aes(
      color = Strain,
      fill  = ifelse(Treatment == "Control", "Control", Strain),
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +
  
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Control" = 21, "Exposed" = 21)
  ) +
  
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = c("white", "black"),
        color = "black"
      )
    )
  ) +
  
  theme_bw() +
  facet_wrap(~ Metric, scales = "free_y") +
  theme(custom_theme) +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
  axis.title.x = element_text(margin = margin(t = -10))
  ) +

  labs(y = "Alpha Diversity Metric Value") +
  
  stat_pvalue_manual(
    alpha_strain_pval_df,
    label = "p.label",
    y.position = c(3.5, 100, 105),
    size = 1.5,
    label.size = 2.5,
    bracket.size = 0.5,
    hide.ns = TRUE
  )
ggsave(file = "strain_alpha_boxplot.svg", plot = strain_alpha_boxplot,width = 4.5, height =3.5, path=path)


ps_control_long <- ps %>%
  ps_filter(Treatment == "Control") %>%
  ps_calc_diversity(rank = "Genus", index = "shannon") %>%
  ps_calc_richness(rank = "Genus", index = "observed") %>%
  samdat_tbl() %>%
  pivot_longer(cols = c("shannon_Genus", "observed_Genus"), names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = recode(Metric, 
                         "shannon_Genus" = "Shannon Entropy", 
                         "observed_Genus" = "# Observed Genera"))

#Alpha Diversity plots For Just Control Samples
# Figure 3A
strain_alpha_control_boxplot <- ps_control_long %>%
  ggplot(aes(y = Value, x = Strain)) +
  
  # Boxplot outlines by strain
  geom_boxplot(
    width = 0.7,
    aes(color = Strain),
    outlier.shape = NA
  ) +
  
  # Jittered points: outline = strain, fill = white (control), shape = 21
  geom_jitter(
    aes(
      color = Strain,
      shape = Treatment
    ),
    fill = "white",        # all control points unfilled
    size = 1,
    stroke = 0.75
  ) +
  
  # Shapes for control group
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Control" = 21)
  ) +
  
  # Strain outline colors
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  # Legend cleanup
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = "white",
        color = "black"
      )
    )
  ) +
  
  theme_bw() +
  facet_wrap(~ Metric, scales = "free_y") +
  theme(custom_theme) +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
  axis.title.x = element_text(margin = margin(t = -10))
  ) +

  
  labs(y = "Alpha Diversity Metric Value") +
  
  # Significance bars
  stat_pvalue_manual(
    alpha_control_pval_df,
    label = "p.label",
    y.position = c(
      3.7, 4.15, 4.75,   # Shannon panel
      3.5, 4.5, 3.85,    # Shannon panel additional?
      100, 110, 126,     # Observed Genera panel
      95, 120, 105       # Observed Genera panel additional?
    ),
    size = 1.5,
    label.size = 2.5,
    bracket.size = 0.5,
    hide.ns = FALSE
  )
ggsave(file = "strain_alpha_control_boxplot.svg", plot = strain_alpha_control_boxplot,width = 4.5, height =3.5, path=path)


#Now for the Exposed Samples only
ps_exposed_long <- ps %>%
  ps_filter(Treatment == "Exposed") %>%
  ps_calc_diversity(rank = "Genus", index = "shannon") %>%
  ps_calc_richness(rank = "Genus", index = "observed") %>%
  samdat_tbl() %>%
  pivot_longer(cols = c("shannon_Genus", "observed_Genus"), names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = recode(Metric, 
                         "shannon_Genus" = "Shannon Entropy", 
                         "observed_Genus" = "# Observed Genera"))


#Alpha Diversity plots For Just exposed Samples
# Figure 3B
strain_alpha_exposed_boxplot <- ps_exposed_long %>%
  ggplot(aes(y = Value, x = Strain)) +
  
  # Boxplot outlines by strain
  geom_boxplot(
    width = 0.7,
    aes(color = Strain),
    outlier.shape = NA
  ) +
  
  # Jittered points: outline = strain, fill = strain color, shape = 21
  geom_jitter(
    aes(
      color = Strain,
      fill  = Strain,       # all exposed samples filled by strain
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75
  ) +
  
  # Fill colors for exposed samples
  scale_fill_manual(
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  # Shapes for exposed group
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Exposed" = 21)
  ) +
  
  # Strain outline colors
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  # Legend cleanup
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = "black",   # exposed = filled
        color = "black"
      )
    )
  ) +
  
  theme_bw() +
  facet_wrap(~ Metric, scales = "free_y") +
  theme(custom_theme) +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
  axis.title.x = element_text(margin = margin(t = -10))
  ) +
  
  labs(y = "Alpha Diversity Metric Value") +
  
  # Significance bars
  stat_pvalue_manual(
    alpha_exposed_pval_df,
    label = "p.label",
    y.position = c(
      3.7, 4.15, 4.75,   # Shannon panel
      3.5, 4.5, 3.85,    # Shannon panel additional
      100, 110, 126,     # Observed Genera panel
      95, 120, 105       # Observed Genera panel additional
    ),
    size = 1.5,
    label.size = 2.5,
    bracket.size = 0.5,
    hide.ns = FALSE
  )
ggsave(file = "strain_alpha_exposed_boxplot.svg", plot = strain_alpha_exposed_boxplot,width = 4.5, height = 3.5, path=path)

#facet wrapped boxplots for treatments showing both alpha metrics
# Figure 4A
treatment_boxplot <- ggplot(ps_long, aes(x = Treatment, y = Value)) +
  
  geom_boxplot(
    width = 0.7,
    outlier.shape = NA
  ) +
  
  geom_jitter(
    aes(
      color = Strain,
      fill  = ifelse(Treatment == "Control", "Control", Strain),
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75,
    show.legend = TRUE
  ) +
  
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +
  
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Control" = 21, "Exposed" = 21)
  ) +
  
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = c("white", "black"),
        color = "black"
      )
    )
  ) +
  
  theme_bw() +
  facet_wrap(~ Metric, scales = "free_y") +
  theme(custom_theme) +
  labs(y = "Alpha Diversity Metric Value", x = "Exposure Group") +
  
  stat_pvalue_manual(
    alpha_treatment_pval_df,
    label = "p.label",
    y.position = c(3.75, 100),
    size = 1.5,
    label.size = 2.5,
    bracket.size = 0.5,
    hide.ns = FALSE
  )

ggsave(file = "treatment_boxplot.svg", plot = treatment_boxplot,width = 4.5, height =3.5, path=path)

#Alpha Diversity by Worm Burden
# Figure S2A
shannon_worm_plot <- ps %>%
  ps_calc_diversity(rank = "Genus", index = "shannon") %>%
  ps_calc_richness(rank = "Genus", index = "observed") %>%
  samdat_tbl() %>%
  ggplot(aes(x = shannon_Genus, y = Total.Worm.Count)) +
  
  geom_point(
    aes(
      color = Strain,
      fill  = ifelse(Treatment == "Control", "Control", Strain),
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75
  ) +
  
  scale_color_manual(
    name = "Strain",
    values = strain_colors
  ) +
  
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +
  
  scale_shape_manual(
    name = "Exposure Group",
    values = shape_values
  ) +
  
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(override.aes = list(fill = c("white", "black")))
  ) +
  
  theme(custom_theme) +
  labs(x = "Shannon Entropy", y = "Total Worm Count") +
  theme_bw(base_size = 12)
ggsave(file = "shannon_worm_plot.svg", plot = shannon_worm_plot,width = 4.75, height =3, path=path)

#Alpha Diversity by Worm Burden
# Figure S2B
observed_worm_plot <- ps %>%
  ps_calc_diversity(rank = "Genus", index = "shannon") %>%
  ps_calc_richness(rank = "Genus", index = "observed") %>%
  samdat_tbl() %>%
  ggplot(aes(x = observed_Genus, y = Total.Worm.Count)) +
  
  geom_point(
    aes(
      color = Strain,
      fill  = ifelse(Treatment == "Control", "Control", Strain),
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75
  ) +
  
  scale_color_manual(
    name = "Strain",
    values = strain_colors
  ) +
  
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +
  
  scale_shape_manual(
    name = "Exposure Group",
    values = shape_values
  ) +
  
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(override.aes = list(fill = c("white", "black")))
  ) +
  
  theme(custom_theme) +
  labs(x = "Observed Richness", y = "Total Worm Count") +
  theme_bw(base_size = 12)

ggsave(file = "observed_worm_plot.svg", plot = observed_worm_plot,width = 4.75, height =3, path=path)

```

#Beta Diversity
```{r}
# PCoA with just control samples
# Figure 3D
pcoa_control <- ps %>%
  ps_filter(Treatment == "Control") %>%
  tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
  tax_agg(rank = "Genus") %>%
  dist_calc(dist = "bray") %>%
  ord_calc(method = "PCoA") %>%
  ord_plot(alpha = 0) +   # remove ghost points
  
  geom_point(
    aes(color = Strain, shape = Treatment),
    fill = "white",
    size = 1,
    stroke = 0.75
  ) +
  
  scale_color_manual(values = strain_colors) +
  scale_shape_manual(values = c("Control" = 21)) +
  
  labs(shape = "Exposure Group") +
  
  stat_ellipse(aes(color = Strain)) +
  
  coord_fixed(0.9) +
  theme_bw() +
  theme(custom_theme)
ggsave(file = "pcoa_control.svg", , plot = pcoa_control, height = 3.75, width = 5, path = path)

#Cropped PCoA plot
# not included in the manuscript
pcoa_all <- ps %>%
  tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
  tax_agg(rank = "Genus") %>%
  dist_calc(dist = "bray") %>%
  ord_calc(method = "PCoA") %>%
  ord_plot(alpha = 0) +   # <-- remove ghost points
  
  geom_point(
    aes(
      color = Strain,
      fill  = ifelse(Treatment == "Control", "Control", as.character(Strain)),
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +
  
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Control" = 21, "Exposed" = 21)
  ) +
  
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = c("white", "black"),
        color = "black"
      )
    )
  ) +
  
  scale_x_continuous(limits = c(-1.5, 1)) +
  scale_y_continuous(limits = c(-2, 2)) +
  
  stat_ellipse(aes(color = Strain)) +
  
  theme_bw() +
  coord_fixed(0.7) +
  theme(custom_theme)
ggsave(file = "pcoa_all.svg", , plot = pcoa_all, height = 3.5, width = 3.5, path = path)

#PCoA plot with ellipses for Treatment dispersion
# Figure 3C
pcoa_all_treatment <- ps %>%
  tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
  tax_agg(rank = "Genus") %>%
  dist_calc(dist = "bray") %>%
  ord_calc(method = "PCoA") %>%
  ord_plot(alpha = 0) +   # <-- remove ghost points
  
  geom_point(
    aes(
      color = Strain,
      fill  = ifelse(Treatment == "Control", "Control", as.character(Strain)),
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +
  
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Control" = 21, "Exposed" = 21)
  ) +
  
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = c("white", "black"),
        color = "black"
      )
    )
  ) +
  
  stat_ellipse(aes(linetype = Treatment)) +
  
  
scale_linetype_manual(
  name = "Exposure Group",
  values = c("Control" = "dashed", "Exposed" = "solid")
) +

  
  theme_bw() +
  coord_fixed(0.7) +
  theme(custom_theme)
ggsave(file = "pcoa_all_treatment.svg", , plot = pcoa_all_treatment, height = 3.75, width = 5, path = path)

#PCoA plot now for the supplemental
# Figure S3
supplemental_pcoa_all <- pcoa_all <- ps %>%
  tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
  tax_agg(rank = "Genus") %>%
  dist_calc(dist = "bray") %>%
  ord_calc(method = "PCoA") %>%
  ord_plot(alpha = 0) +   # <-- remove ghost points
  
  geom_point(
    aes(
      color = Strain,
      fill  = ifelse(Treatment == "Control", "Control", as.character(Strain)),
      shape = Treatment
    ),
    size = 1,
    stroke = 0.75,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +
  
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  scale_shape_manual(
    name = "Exposure Group",
    values = c("Control" = 21, "Exposed" = 21)
  ) +
  
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(shape = 16, size = 2)),
    shape = guide_legend(
      override.aes = list(
        size  = 2,
        shape = 21,
        fill  = c("white", "black"),
        color = "black"
      )
    )
  ) +
  
  stat_ellipse(aes(color = Strain)) +
  
  theme_bw() +  
  coord_fixed(0.9) +
  theme(custom_theme)
ggsave(file = "supplemental_pcoa_all.svg", , plot = supplemental_pcoa_all, height = 4, width = 5, path = path)

#without strain colors, with fill by worm burden
# Figure 4C
pcoa_wormburden <- ps %>%
  tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
  tax_agg(rank = "Genus") %>%
  dist_calc(dist = "bray") %>%
  ord_calc(method = "PCoA") %>%
  ord_plot(alpha = 0.7) +

  # Points filled by worm burden, shaped by treatment
  geom_point(
    aes(
      fill = Total.Worm.Count,
      shape = Treatment
    ),
    color = "black",
    size = 2,
    stroke = 0.3
  ) +

  # Continuous fill scale with 0 = white
  scale_fill_gradient(
    name = "Worm Burden",
    low = "white",
    high = "#1a1a1a"
  ) +

  # Shapes for treatment (but legend removed below)
  scale_shape_manual(
    values = c("Control" = 21, "Exposed" = 21)
  ) +

  # Remove exposure-group legend
  guides(
    shape = "none"
  ) +

 # scale_x_continuous(limits = c(-1.5, 1)) +
 # scale_y_continuous(limits = c(-2, 2)) +

  theme_classic(12) +
  coord_fixed(0.7) +
  theme_bw() +
  theme(custom_theme) #+
  #theme(legend.direction = "horizontal")
ggsave(file = "pcoa_wormburden.svg", , plot = pcoa_wormburden, height = 3.5, width = 5, path = path)

#add DPE shapes
# Figure 4B
pcoa_dpe <- ps %>%
  tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
  tax_agg(rank = "Genus") %>%
  dist_calc(dist = "bray") %>%
  ord_calc(method = "PCoA") %>%
  ord_plot(alpha = 0) +   # remove ghost points
  
  geom_point(
    aes(
      color = Strain,
      fill  = ifelse(Treatment == "Control", "Control", as.character(Strain)),
      shape = DPE,
      alpha = Treatment        # <-- restores Exposure Group legend
    ),
    size = 1,
    stroke = 0.75,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(
    name = "Fill",
    values = c(
      "Control" = "white",
      "5D"      = "#E63073",
      "AB"      = "#82B7E6",
      "Casper"  = "#FFC107",
      "TU"      = "#44B5A2"
    )
  ) +
  
  scale_color_manual(
    name = "Strain",
    values = c(
      "5D"     = "#E63073",
      "AB"     = "#82B7E6",
      "Casper" = "#FFC107",
      "TU"     = "#44B5A2"
    )
  ) +
  
  scale_shape_manual(
    name = "DPE",
    values = c("31" = 21, "44" = 24)
  ) +
  
  scale_alpha_manual(
    name = "Exposure Group",
    values = c("Control" = 1, "Exposed" = 1)
  ) +
  
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 2)),
    shape = guide_legend(override.aes = list(size = 2)),
    alpha = guide_legend(
      override.aes = list(
        shape = 21,
        fill  = c("white", "black"),
        color = "black",
        size  = 2
      )
    )
  ) +
  
  coord_fixed(0.7) +
  theme_bw() +
  theme(custom_theme)

ggsave(file = "pcoa_dpe.svg", , plot = pcoa_dpe, height = 3.5, width = 5, path = path)
```
#Taxa
```{r}
#all_sig_results dataframe created in the MaAsLin2 Rmd

#format the names
all_sig_results <- all_sig_results %>%
  mutate(feature = gsub("\\.", " ", feature))

# modify taxa names of interest
all_sig_results <- all_sig_results %>%
  mutate(feature = ifelse(feature %in% c("Barnesiellaceae.Genus", "Cetobacterium", "Paraclostridium" , "Pseudomonas", "ZOR0006"), 
                          paste0(feature, "*"), 
                          feature))

# Get unique features in the order they appear on the x-axis
x_labels <- unique(all_sig_results$feature)

# Create a named vector of HTML-styled labels
label_colors <- sapply(x_labels, function(label) {
  if (grepl("\\*", label)) {
    paste0("<span style='color:red;'>", label, "</span>")
  } else {
    paste0("<span style='color:black;'>", label, "</span>")
  }
})
names(label_colors) <- x_labels

# Create the heatmap
heatmap_plot <- ggplot(all_sig_results, aes(x = feature, y = Strain, fill = coef)) +
  geom_tile() +
  geom_text(aes(label = ifelse(qval < 0.1, "*", "")), size = 3, color = "black") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +
  scale_x_discrete(labels = label_colors) +
  theme_minimal() +
  theme(
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = -17)),
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    legend.position = "bottom",
    legend.justification = c(1, 0),
    legend.box.just = "right",
    legend.margin = margin(t = -45),
    panel.grid.major = element_line(color = "white"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_rect(color = "gray95"),
    plot.background = element_rect(fill = "white"),
  ) +
  labs(x = "Taxa", y = "Strain")
# Save the plot
ggsave("diff_assoc_heatmap_wide.svg", plot = heatmap_plot, height = 4, width = 7.5, units = "in", path = path)

# Tall orientation
# Get unique features in the order they appear on the y-axis
y_labels <- unique(all_sig_results$feature)

# Create a named vector of HTML-styled labels
label_colors <- sapply(y_labels, function(label) {
  if (grepl("\\*", label)) {
    paste0("<span style='color:red;'>", label, "</span>")
  } else {
    paste0("<span style='color:black;'>", label, "</span>")
  }
})
names(label_colors) <- y_labels

heatmap_plot_tall <- ggplot(all_sig_results, aes(x = Strain, y = feature, fill = coef)) +
  geom_tile() +
  geom_text(aes(label = ifelse(qval < 0.1, "*", "")), size = 3, color = "black") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +
  scale_y_discrete(labels = label_colors, limits = rev(y_labels)) +
  theme_minimal() +
  theme(
    axis.text.y = element_markdown(hjust = 1, vjust = 0.6),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 15)),
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    legend.position = "bottom",
    legend.justification = c(-0.5, 0),
    legend.box.just = "right",
    legend.margin = margin(t = -35),
    panel.grid.major = element_line(color = "white"),
    panel.grid.minor = element_line(color = "gray95"),
    panel.background = element_rect(color = "gray95"),
    plot.background = element_rect(fill = "white"),
  ) +
  labs(x = "Strain", y = "Taxa")

# Save the plot
ggsave("diff_assoc_heatmap_tall.svg", plot = heatmap_plot_tall, height = 7.5, width = 7.5, units = "in", path = path)

# Create the table in long format with coefficients in the cells
tax_table <- as.data.frame(ps@tax_table)

# Clean the feature column if it contains asterisks
all_sig_results_tax <- all_sig_results %>%
  mutate(feature_clean = gsub("\\*", "", feature))

# Create a named vector from tax_table: names are genera, values are families
genus_to_family <- tax_table$Family
names(genus_to_family) <- tax_table$Genus

genus_to_order <- tax_table$Order
names(genus_to_order) <- tax_table$Genus

genus_to_class <- tax_table$Class
names(genus_to_class) <- tax_table$Genus

# Map the family to each feature
all_sig_results_tax <- all_sig_results_tax %>%
  mutate(Family = genus_to_family[feature_clean]) %>%
  mutate(Order = genus_to_order[feature_clean]) %>%
    mutate(Class = genus_to_class[feature_clean])

# add faceting
heatmap_plot_tax <- ggplot(all_sig_results_tax, aes(x = Strain, y = feature, fill = coef)) +
  geom_tile() +
  geom_text(aes(label = round(coef, 2)), size = 2.5, color = "black") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +
  scale_y_discrete(labels = label_colors) +
  facet_grid(Class ~ ., scales = "free_y", space = "free_y") +
  theme_minimal() +
  theme(
    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 1,  face = "bold", size = 8),  # Left-align strip text
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 0.5),
    axis.text.y = element_markdown(),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 15)),
    axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 45)),
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    legend.position = "bottom",
    legend.justification = c(-1.5, 0),
    legend.box.just = "right",
    legend.margin = margin(t = -45),
    panel.grid.major = element_line(color = "white"),
    panel.grid.minor = element_line(color = "gray90"),
    panel.background = element_rect(color = "gray95"),
    plot.background = element_rect(fill = "white")
  ) +
  labs(x = "Strain", y = "Taxa")


# Save the plot
# Figure 5
ggsave("diff_assoc_heatmap_tax.svg", plot = heatmap_plot_tax, height = 7.5, width = 6.5, units = "in", path = path)



# MaAsLin Output for Strain associated taxa in just the control fish
# Only control fish, Genus level, just Strain
fit_data_control_genus_strain = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_control_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_control_genus_strain", 
                    fixed_effects  = c("Strain"))
View(fit_data_control_genus_strain$results)


# Extract the dataframe
df <- fit_data_control_genus_strain$results

# Filter for qval < 0.25
filtered_df <- df %>%
  filter(qval < 0.25)

# Create the transformed value
filtered_df <- filtered_df %>%
  mutate(transformed_value = -log(qval) * sign(coef))

# Create the heatmap
heatmap_strain_taxa <- ggplot(
  filtered_df,
  aes(x = value, y = feature, fill = transformed_value)
) +
  geom_tile(color = "white") +

  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "-log(qval) × sign(coef)",
    guide = guide_colorbar(
      direction = "horizontal",
      title.position = "top",
      title.hjust = 0.5,
      barwidth = 12,
      barheight = 0.5
    )
  ) +

  theme_minimal() +
  theme(
    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 1, face = "bold"),
    axis.text.x = element_markdown(angle = 45, hjust = 1),
    axis.text.y = element_markdown(),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 5)),
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.justification = c(1.6, 0),
    legend.margin = margin(t = -55),
    panel.grid.major = element_line(color = "white"),
    panel.grid.minor = element_line(color = "gray90"),
    panel.background = element_rect(color = "gray95"),
    plot.background = element_rect(fill = "white")
  ) +

  labs(
    x = "Strain",
    y = "Taxa"
  )


# Save the plot
# Figure S1B
ggsave("heatmap_strain_taxa.svg", plot = heatmap_strain_taxa, height = 3.5, width = 3.5, units = "in", path = path)
```
```