---
title: "Maaslin2 Analysis"
author: "Emilee Lance"
date: "2024-06-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Maaslin2)
library(microViz)
library(dplyr)
library(ggplot2)
```

```{r set-up}
getwd()
setwd("C:/Users/emile/OneDrive - Oregon State University/Sharpton Lab/Interstrain_Zebrafish_Gut_Microbiome/code")
#Pull files from the ps_object
ps <- readRDS("ps", file = "C:/Users/emile/OneDrive - Oregon State University/Sharpton Lab/Interstrain_Zebrafish_Gut_Microbiome/Data/ps_object.rds")
#Metadata
df_input_metadata <- as.data.frame(as.matrix(ps@sam_data))
df_input_metadata$"Strain_Treatment" <- paste(df_input_metadata$Strain, df_input_metadata$Treatment, sep = "_") #create interaction term
df_input_metadata$Total.Worm.Count <- as.numeric(df_input_metadata$Total.Worm.Count) #convert worm count to numeric
#Set references for all variables
df_input_metadata$Strain = factor(df_input_metadata$Strain,
                                              levels = c("AB", "5D", "Casper", "TU"))
df_input_metadata$DPE = factor(df_input_metadata$DPE,
                                              levels = c("31", "44"))
df_input_metadata$Pathology.Results = factor(df_input_metadata$Pathology.Results,
                                              levels = c("negative", "positive"))
df_input_metadata$Strain_Treatment = factor(df_input_metadata$Strain_Treatment,
                                              levels = c("AB_Control", "AB_Exposed", "5D_Control", "5D_Exposed", 
                                                         "Casper_Control", "Casper_Exposed", "TU_Control", "TU_Exposed"))
View(df_input_metadata)
#subset metadata
df_control_metadata <- subset(df_input_metadata, Treatment == "Control")
df_exposed_metadata <- subset(df_input_metadata, Treatment == "Exposed")
df_positive_metadata <- subset(df_input_metadata, Pathology.Results == "positive")
df_5D_metadata <- subset(df_input_metadata, Strain =="5D")
df_5D_control_metadata <- subset(df_5D_metadata, Treatment == "Control")
df_5D_exposed_metadata <- subset(df_5D_metadata, Treatment == "Exposed")
df_5D_31_metadata <- subset(df_5D_metadata, DPE == "31")
df_5D_44_metadata <- subset(df_5D_metadata, DPE == "44")
df_AB_metadata <- subset(df_input_metadata, Strain == "AB")
df_AB_control_metadata <- subset(df_AB_metadata, Treatment == "Control")
df_AB_exposed_metadata <- subset(df_AB_metadata, Treatment == "Exposed")
df_AB_31_metadata <- subset(df_AB_metadata, DPE == "31")
df_AB_44_metadata <- subset(df_AB_metadata, DPE == "44")
df_Casper_metadata <- subset(df_input_metadata, Strain =="Casper")
df_Casper_control_metadata <- subset(df_Casper_metadata, Treatment == "Control")
df_Casper_exposed_metadata <- subset(df_Casper_metadata, Treatment == "Exposed")
df_Casper_31_metadata <- subset(df_Casper_metadata, DPE == "31")
df_Casper_44_metadata <- subset(df_Casper_metadata, DPE == "44")
df_TU_metadata <- subset(df_input_metadata, Strain =="TU")
df_TU_control_metadata <- subset(df_TU_metadata, Treatment == "Control")
df_TU_exposed_metadata <- subset(df_TU_metadata, Treatment == "Exposed")
df_TU_31_metadata <- subset(df_TU_metadata, DPE == "31")
df_TU_44_metadata <- subset(df_TU_metadata, DPE == "44")

#Species level OTU
ps <- ps %>%
     tax_agg(rank = "Species")
df_input_data_species <- as.data.frame(as.matrix(ps@otu_table))
View(df_input_data_species)
#Genus level OTUs
ps <- ps %>%
     tax_agg(rank = "Genus")
df_input_data_genus <- as.data.frame(as.matrix(ps@otu_table))

```

# Run MaAsLin2 on all fish, species level
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_species = Maaslin2(input_data = df_input_data_species, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_species", 
                    fixed_effects  = c("Strain", "Treatment", "Strain_Treatment", "Total.Worm.Count"))
View(fit_data_all_species$results)
```

#All fish, Genus level, Strain + Treatment + Worm Count
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_genus = Maaslin2(input_data = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_genus", 
                    fixed_effects  = c("Strain", "Treatment", "Strain_Treatment", "Total.Worm.Count"))
View(fit_data_all_genus$results)
```

#All fish, Genus level, Strain
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_genus_strain = Maaslin2(input_data = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_genus_strain", 
                    fixed_effects  = c("Strain"))
View(fit_data_all_genus_strain$results)
```

#All fish, Genus level, Strain
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_genus_strain_dpe = Maaslin2(input_data = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_genus_strain_dpe", 
                    fixed_effects  = c("Strain", "DPE"))
View(fit_data_all_genus_strain_dpe$results)
```

#All fish, Genus level, no worms
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_genus_noworms = Maaslin2(input_data = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_genus_noworms", 
                    fixed_effects  = c("Strain", "Treatment", "Strain_Treatment"))
View(fit_data_all_genus_noworms$results)
```

#All fish, Genus level, worms (presence and association)
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_genus_worms = Maaslin2(input_data = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_genus_worms", 
                    fixed_effects  = c("Pathology.Results", "Total.Worm.Count"))
View(fit_data_genus_worms$results)

#create new heatmap
sig_worms <- subset(fit_data_genus_worms$results, qval < 0.25)
#long format
worms_heatmap_plot <- ggplot(sig_worms, aes(x = metadata, y = feature, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
   ) +
  labs(x = "Worm Association", y = "Taxa", title = "Taxa Associated with Worm Presence and Burden")
# Save the plot
ggsave("worms_heatmap_plot_long.png", plot = worms_heatmap_plot, height = 6, width = 6, units = "in")

# Wide format
worms_heatmap_plot <- ggplot(sig_worms, aes(x = feature, y = metadata, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1)) +
  labs(x = "Taxa", y = "Worm Associations", title = "Taxa Associated with Worm Presence and Burden")
# Save the plot
ggsave("worms_heatmap_plot_wide_0.05.png", plot = worms_heatmap_plot, height = 6, width = 16, units = "in")

#extract names of significantly associated taxa
sig_taxa <- subset(fit_data_genus_worms$results, qval < 0.25) %>%
  subset(metadata == "Pathology.Results")
sig_taxa_names <- as.list(sig_taxa$feature)
#create new input dataframe
df_positive_input_data_genus <- df_input_data_genus[, colnames(df_input_data_genus) %in% sig_taxa_names]
```

#All fish, Genus level, worms (association)
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_genus_count = Maaslin2(input_data = df_positive_input_data_genus, 
                    input_metadata = df_positive_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_genus_count", 
                    fixed_effects  = c("Total.Worm.Count"))
View(fit_data_genus_count$results)
```

# All fish, only interaction, species level
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_worm_species = Maaslin2(input_data = df_input_data_species, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_worm_species", 
                    fixed_effects  = c("Strain_Treatment", "Total.Worm.Count"))
View(fit_data_all_worm_species$results)
```

#All fish, Genus level, no interaction
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_genus_noint = Maaslin2(input_data = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_genus_noint", 
                    fixed_effects  = c("Strain", "Treatment"))
View(fit_data_all_genus_noint$results)
```

#All fish, Genus level, Treatment
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_genus_treat = Maaslin2(input_data = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_genus_treat", 
                    fixed_effects  = c("Treatment"))
View(fit_data_all_genus_treat$results)
```

#All fish, Genus level, Treatment and DPE
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_all_genus_treat_dpe = Maaslin2(input_data = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_genus_treat_dpe", 
                    fixed_effects  = c("DPE", "Treatment"))
View(fit_data_all_genus_treat_dpe$results)
```

# With only Exposed Fish, species level
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_exposed_species = Maaslin2(input_data     = df_input_data_species, 
                    input_metadata = df_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_exposed_species", 
                    fixed_effects  = c("Strain","Total.Worm.Count", "DPE"))
View(fit_data_exposed_species$results)
```

# Exposed Fish, Genus Level
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_exposed_genus = Maaslin2(input_data   = df_input_data_genus, 
                    input_metadata = df_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_exposed_genus", 
                    fixed_effects  = c("Strain","Total.Worm.Count", "DPE"))
View(fit_data_exposed_genus$results)
```

# Exposed Fish, Genus Level, Pathology
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_exposed_genus_pathology = Maaslin2(input_data   = df_input_data_genus, 
                    input_metadata = df_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_exposed_genus_pathology", 
                    fixed_effects  = c("Strain","Pathology.Results", "DPE"))
View(fit_data_exposed_genus_pathology$results)
```

# With only Exposed Fish, species level, no worms
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_exposed_species_noworms = Maaslin2(input_data     = df_input_data_species, 
                    input_metadata = df_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_exposed_species_noworms", 
                    fixed_effects  = c("Strain", "DPE"))
View(fit_data_exposed_species_noworms$results)
```

# Exposed Fish, Genus Level, no worms
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_exposed_genus_noworms = Maaslin2(input_data   = df_input_data_genus, 
                    input_metadata = df_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_exposed_genus_noworms", 
                    fixed_effects  = c("Strain", "DPE"))
View(fit_data_exposed_genus_noworms$results)
```

# With only Exposed Fish, species level, only worms
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_exposed_species_onlyworms = Maaslin2(input_data     = df_input_data_species, 
                    input_metadata = df_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_exposed_species_onlyworms", 
                    fixed_effects  = c("Strain", "Total.Worm.Count"))
View(fit_data_exposed_species_onlyworms$results)
```

# Exposed Fish, Genus Level, only worms
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_exposed_genus_onlyworms = Maaslin2(input_data   = df_input_data_genus, 
                    input_metadata = df_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_exposed_genus_onlyworms", 
                    fixed_effects  = c("Strain", "Total.Worm.Count"))
View(fit_data_exposed_genus_onlyworms$results)
```

# Only Control Fish, species level
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_control_species = Maaslin2(input_data     = df_input_data_species, 
                    input_metadata = df_control_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_control_species", 
                    fixed_effects  = c("Strain", "DPE"))
View(fit_data_control_species$results)
```

# Only control fish, Genus level
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_control_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_control_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_control_genus", 
                    fixed_effects  = c("Strain", "DPE"))
View(fit_data_control_genus$results)
```


# Only control fish, Genus level, just Strain
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
fit_data_control_genus_strain = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_control_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_control_genus_strain", 
                    fixed_effects  = c("Strain"))
View(fit_data_control_genus_strain$results)
```

# All fish, each strain separate by treatment
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
#5D
fit_data_5D_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_5D_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_5D_genus", 
                    fixed_effects  = c("Treatment"))

#AB
fit_data_AB_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_AB_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_AB_genus", 
                    fixed_effects  = c("Treatment"))

#Casper
fit_data_Casper_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_Casper_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_Casper_genus", 
                    fixed_effects  = c("Treatment"))

#TU
fit_data_TU_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_TU_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_TU_genus", 
                    fixed_effects  = c("Treatment"))
```
#All fish, Separate strains, compile results
```{r}
# add strain back to dataframe and pull significant results
fit_data_5D_genus$results$Strain <- "5D"
fit_data_5D_genus$sig_results <- subset(fit_data_5D_genus$results, qval < 0.25)

fit_data_AB_genus$results$Strain <- "AB"
fit_data_AB_genus$sig_results <- subset(fit_data_AB_genus$results, qval < 0.25)

fit_data_Casper_genus$results$Strain <- "Casper"
fit_data_Casper_genus$sig_results <- subset(fit_data_Casper_genus$results, qval < 0.25)

fit_data_TU_genus$results$Strain <- "TU"
fit_data_TU_genus$sig_results <- subset(fit_data_TU_genus$results, qval < 0.25)

#combine dataframes
all_sig_results <- bind_rows(fit_data_5D_genus$sig_results, fit_data_AB_genus$sig_results, fit_data_Casper_genus$sig_results,fit_data_TU_genus$sig_results)
#add column denoting significance
#all_sig_results$significance <- ifelse(all_sig_results$qval)
#View(all_sig_results)

#count occurrences significant associations by strain
all_sig_results %>%
group_by(Strain) %>%
  summarise(
    Positive = sum(coef> 0),
    Negative = sum(coef < 0)
  )

library(reshape2)

# Convert to wide format for heatmap
heatmap_data <- dcast(all_sig_results, feature ~ Strain, value.var = "coef")
rownames(heatmap_data) <- heatmap_data$feature
heatmap_data <- heatmap_data[, -1]  # Remove first column (feature)


library(gplots)
# Create heatmap
taxa_heatmap <- heatmap.2(as.matrix(heatmap_data),
          Rowv = FALSE, Colv = FALSE,      # Turn off row and column clustering
          dendrogram = "none",             # Turn off dendrogram
          trace = "none",                  # Turn off trace lines
          margins = c(5, 20),              # Adjust margins
          main = "(−log(qval)*sign(coeff))", # Title of the heatmap
          key = TRUE,                      # Show color key
          keysize = 1.0,                   # Size of color key
          density.info = "none",           # Turn off density plot
          symkey = TRUE,                   # Show symmetric key
          col = colorRampPalette(c("red", "white", "blue")),        # Color palette
          scale = "row",                   # Scale rows (optional)
          cexCol = 0.8, cexRow = 0.8,      # Adjust text size
          labRow = heatmap_data$feature,   # Row labels
          labCol = colnames(heatmap_data)  # Column labels
)

#Using ggplot instead
heatmap_plot <- ggplot(all_sig_results, aes(x = Strain, y = feature, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white")) +  # Optional: adjust theme
  labs(x = "Strain", y = "Taxa", title = "Differentially Associated Taxa by Strain")  # Optional: add labels and title
# Save the plot
ggsave("diff_assoc_heatmap.png", plot = heatmap_plot, height = 10, width = 8, units = "in")

# rotated for PPT
heatmap_plot <- ggplot(all_sig_results, aes(x = feature, y = Strain, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1)) +
  labs(x = "Taxa", y = "Strain", title = "Differentially Associated Taxa by Strain")
# Save the plot
ggsave("diff_assoc_heatmap_wide.png", plot = heatmap_plot, height = 6, width = 16, units = "in", path = path)
```

# Control fish, each strain separate by DPE
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
#Controls
#5D
fit_data_5D_DPE_control_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_5D_control_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_5D_DPE_control_genus", 
                    fixed_effects  = c("DPE"))

#AB
fit_data_AB_DPE_control_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_AB_control_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_AB_DPE_control_genus", 
                    fixed_effects  = c("DPE"))

#Casper
fit_data_Casper_DPE_control_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_Casper_control_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_Casper_DPE_control_genus", 
                    fixed_effects  = c("DPE"))

#TU
fit_data_TU_DPE_control_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_TU_control_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_TU_DPE_control_genus", 
                    fixed_effects  = c("DPE"))
```
#Control fish, Separate strains by DPE, compile results
```{r}
# add strain back to dataframe and pull significant results
fit_data_5D_DPE_control_genus$results$Strain <- "5D"
fit_data_5D_DPE_control_genus$sig_results <- subset(fit_data_5D_DPE_control_genus$results, qval < 0.25)

fit_data_AB_DPE_control_genus$results$Strain <- "AB"
fit_data_AB_DPE_control_genus$sig_results <- subset(fit_data_AB_DPE_control_genus$results, qval < 0.25)

fit_data_Casper_DPE_control_genus$results$Strain <- "Casper"
fit_data_Casper_DPE_control_genus$sig_results <- subset(fit_data_Casper_DPE_control_genus$results, qval < 0.25)

fit_data_TU_DPE_control_genus$results$Strain <- "TU"
fit_data_TU_DPE_control_genus$sig_results <- subset(fit_data_TU_DPE_control_genus$results, qval < 0.25)

#combine dataframes
DPE_control_sig_results <- bind_rows(fit_data_5D_DPE_control_genus$sig_results, fit_data_AB_DPE_control_genus$sig_results, fit_data_Casper_DPE_control_genus$sig_results,fit_data_TU_DPE_control_genus$sig_results)
#add column denoting significance
DPE_control_sig_results$significance <- ifelse(DPE_control_sig_results$qval) # haven't figured this out yet
View(DPE_control_sig_results)

#count occurrences significant associations by strain
DPE_control_sig_results %>%
group_by(Strain) %>%
  summarise(
    Positive = sum(coef> 0),
    Negative = sum(coef < 0)
  )

library(reshape2)

# Wide format
DPE_control_heatmap_plot <- ggplot(DPE_control_sig_results, aes(x = feature, y = Strain, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1)) +
  labs(x = "Taxa", y = "Strain", title = "Differentially Associated Taxa by Strain and DPE in Control fish")
# Save the plot
ggsave("diff_assoc_heatmap_DPE_control_wide.png", plot = DPE_control_heatmap_plot, height = 6, width = 16, units = "in")

#long format
DPE_control_heatmap_plot <- ggplot(DPE_control_sig_results, aes(x = Strain, y = feature, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
   ) +
  labs(x = "Strain", y = "Taxa", title = "Differentially Associated Taxa by Strain and DPE in Control fish")
# Save the plot
ggsave("diff_assoc_heatmap_DPE_control_long.png", plot = DPE_control_heatmap_plot, height = 6, width = 6, units = "in")
```













# Exposed fish, Separate strains, split by DPE
```{r}
#Exposed
#5D
fit_data_5D_DPE_exposed_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_5D_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_5D_DPE_exposed_genus", 
                    fixed_effects  = c("DPE"))

#AB
fit_data_AB_DPE_exposed_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_AB_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_AB_DPE_exposed_genus", 
                    fixed_effects  = c("DPE"))

#Casper
fit_data_Casper_DPE_exposed_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_Casper_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_Casper_DPE_exposed_genus", 
                    fixed_effects  = c("DPE"))

#TU
fit_data_TU_DPE_exposed_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_TU_exposed_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_TU_DPE_exposed_genus", 
                    fixed_effects  = c("DPE"))
```
#Exposed fish, Separate strains by DPE, compile results
```{r}
# add strain back to dataframe and pull significant results
fit_data_5D_DPE_exposed_genus$results$Strain <- "5D"
fit_data_5D_DPE_exposed_genus$sig_results <- subset(fit_data_5D_DPE_exposed_genus$results, qval < 0.25)

fit_data_AB_DPE_exposed_genus$results$Strain <- "AB"
fit_data_AB_DPE_exposed_genus$sig_results <- subset(fit_data_AB_DPE_exposed_genus$results, qval < 0.25)

fit_data_Casper_DPE_exposed_genus$results$Strain <- "Casper"
fit_data_Casper_DPE_exposed_genus$sig_results <- subset(fit_data_Casper_DPE_exposed_genus$results, qval < 0.25)

fit_data_TU_DPE_exposed_genus$results$Strain <- "TU"
fit_data_TU_DPE_exposed_genus$sig_results <- subset(fit_data_TU_DPE_exposed_genus$results, qval < 0.25)

#combine dataframes
DPE_exposed_sig_results <- bind_rows(fit_data_5D_DPE_exposed_genus$sig_results, fit_data_AB_DPE_exposed_genus$sig_results, fit_data_Casper_DPE_exposed_genus$sig_results,fit_data_TU_DPE_exposed_genus$sig_results)
#add column denoting significance
DPE_exposed_sig_results$significance <- ifelse(DPE_exposed_sig_results$qval)# havent figured this out yet
View(DPE_exposed_sig_results)

#count occurrences significant associations by strain
DPE_exposed_sig_results %>%
group_by(Strain) %>%
  summarise(
    Positive = sum(coef> 0),
    Negative = sum(coef < 0)
  )

# Wide format
DPE_exposed_heatmap_plot <- ggplot(DPE_exposed_sig_results, aes(x = feature, y = Strain, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1)) +
  labs(x = "Taxa", y = "Strain", title = "Differentially Associated Taxa by Strain and DPE in Exposed fish")
# Save the plot
ggsave("diff_assoc_heatmap_DPE_exposed_wide.png", plot = DPE_exposed_heatmap_plot, height = 6, width = 16, units = "in")

#long format
DPE_exposed_heatmap_plot <- ggplot(DPE_exposed_sig_results, aes(x = Strain, y = feature, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
   ) +
  labs(x = "Strain", y = "Taxa", title = "Differentially Associated Taxa by Strain and DPE in Exposed fish")
# Save the plot
ggsave("diff_assoc_heatmap_DPE_exposed_long.png", plot = DPE_exposed_heatmap_plot, height = 6, width = 6, units = "in")
```









# 31 DPE fish, each strain separate by Treatment
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
#Controls
#5D
fit_data_5D_DPE_31_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_5D_31_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_5D_DPE_31_genus", 
                    fixed_effects  = c("Treatment"))

#AB
fit_data_AB_DPE_31_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_AB_31_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_AB_DPE_31_genus", 
                    fixed_effects  = c("Treatment"))

#Casper
fit_data_Casper_DPE_31_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_Casper_31_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_Casper_DPE_31_genus", 
                    fixed_effects  = c("Treatment"))

#TU
fit_data_TU_DPE_31_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_TU_31_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_TU_DPE_31_genus", 
                    fixed_effects  = c("Treatment"))
```

#31 DPE fish, Separate strains by Treatment, compile results
```{r}
# add strain back to dataframe and pull significant results
fit_data_5D_DPE_31_genus$results$Strain <- "5D"
fit_data_5D_DPE_31_genus$sig_results <- subset(fit_data_5D_DPE_31_genus$results, qval < 0.25)

fit_data_AB_DPE_31_genus$results$Strain <- "AB"
fit_data_AB_DPE_31_genus$sig_results <- subset(fit_data_AB_DPE_31_genus$results, qval < 0.25)

fit_data_Casper_DPE_31_genus$results$Strain <- "Casper"
fit_data_Casper_DPE_31_genus$sig_results <- subset(fit_data_Casper_DPE_31_genus$results, qval < 0.25)

fit_data_TU_DPE_31_genus$results$Strain <- "TU"
fit_data_TU_DPE_31_genus$sig_results <- subset(fit_data_TU_DPE_31_genus$results, qval < 0.25)

#combine dataframes
DPE_31_sig_results <- bind_rows(fit_data_5D_DPE_31_genus$sig_results, fit_data_AB_DPE_31_genus$sig_results, fit_data_Casper_DPE_31_genus$sig_results,fit_data_TU_DPE_31_genus$sig_results)
#add column denoting significance
DPE_31_sig_results$significance <- ifelse(DPE_31_sig_results$qval) # haven't figured this out yet
View(DPE_31_sig_results)

#count occurrences significant associations by strain
DPE_31_sig_results %>%
group_by(Strain) %>%
  summarise(
    Positive = sum(coef> 0),
    Negative = sum(coef < 0)
  )

library(reshape2)

# Wide format
DPE_31_heatmap_plot <- ggplot(DPE_31_sig_results, aes(x = feature, y = Strain, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1)) +
  labs(x = "Taxa", y = "Strain", title = "Differentially Associated Taxa by Strain and Treatment in Fish 31 DPE")
# Save the plot
ggsave("diff_assoc_heatmap_DPE_31_wide.png", plot = DPE_31_heatmap_plot, height = 6, width = 16, units = "in")

#long format
DPE_31_heatmap_plot <- ggplot(DPE_31_sig_results, aes(x = Strain, y = feature, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
   ) +
  labs(x = "Strain", y = "Taxa", title = "Differentially Associated Taxa by Strain and Treatment in Fish 31 DPE")
# Save the plot
ggsave("diff_assoc_heatmap_DPE_31_long.png", plot = DPE_31_heatmap_plot, height = 6, width = 6, units = "in")
```

# 44 DPE fish, each strain separate by Treatment
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
#Controls
#5D
fit_data_5D_DPE_44_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_5D_44_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_5D_DPE_44_genus", 
                    fixed_effects  = c("Treatment"))

#AB
fit_data_AB_DPE_44_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_AB_44_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_AB_DPE_44_genus", 
                    fixed_effects  = c("Treatment"))

#Casper
fit_data_Casper_DPE_44_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_Casper_44_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_Casper_DPE_44_genus", 
                    fixed_effects  = c("Treatment"))

#TU
fit_data_TU_DPE_44_genus = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_TU_44_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_TU_DPE_44_genus", 
                    fixed_effects  = c("Treatment"))
```

#44 DPE fish, Separate strains by Treatment, compile results
```{r}
# add strain back to dataframe and pull significant results
fit_data_5D_DPE_44_genus$results$Strain <- "5D"
fit_data_5D_DPE_44_genus$sig_results <- subset(fit_data_5D_DPE_44_genus$results, qval < 0.25)

fit_data_AB_DPE_44_genus$results$Strain <- "AB"
fit_data_AB_DPE_44_genus$sig_results <- subset(fit_data_AB_DPE_44_genus$results, qval < 0.25)

fit_data_Casper_DPE_44_genus$results$Strain <- "Casper"
fit_data_Casper_DPE_44_genus$sig_results <- subset(fit_data_Casper_DPE_44_genus$results, qval < 0.25)

fit_data_TU_DPE_44_genus$results$Strain <- "TU"
fit_data_TU_DPE_44_genus$sig_results <- subset(fit_data_TU_DPE_44_genus$results, qval < 0.25)

#combine dataframes
DPE_44_sig_results <- bind_rows(fit_data_5D_DPE_44_genus$sig_results, fit_data_AB_DPE_44_genus$sig_results, fit_data_Casper_DPE_44_genus$sig_results,fit_data_TU_DPE_44_genus$sig_results)
#add column denoting significance
DPE_44_sig_results$significance <- ifelse(DPE_44_sig_results$qval) # haven't figured this out yet
View(DPE_44_sig_results)

#count occurrences significant associations by strain
DPE_44_sig_results %>%
group_by(Strain) %>%
  summarise(
    Positive = sum(coef> 0),
    Negative = sum(coef < 0)
  )

library(reshape2)

# Wide format
DPE_44_heatmap_plot <- ggplot(DPE_44_sig_results, aes(x = feature, y = Strain, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1)) +
  labs(x = "Taxa", y = "Strain", title = "Differentially Associated Taxa by Strain and Treatment in Fish 44 DPE")
# Save the plot
ggsave("diff_assoc_heatmap_DPE_44_wide.png", plot = DPE_44_heatmap_plot, height = 6, width = 16, units = "in")

#long format
DPE_44_heatmap_plot <- ggplot(DPE_44_sig_results, aes(x = Strain, y = feature, fill = coef)) +
  geom_tile() +  # Use geom_tile to create the heatmap
  scale_fill_gradient2(low = "blue", high = "red", mid = "white") +  # Adjust color gradient
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white"), # Optional: adjust theme
   ) +
  labs(x = "Strain", y = "Taxa", title = "Differentially Associated Taxa by Strain and Treatment in Fish 44 DPE")
# Save the plot
ggsave("diff_assoc_heatmap_DPE_44_long.png", plot = DPE_44_heatmap_plot, height = 6, width = 6, units = "in")
```






















# All fish, Split by DPE and Strain
```{r, message = FALSE, warning = FALSE, include = FALSE, results = hide}
df_input_metadata$"Strain_Treatment_DPE" <- paste(df_input_metadata$Strain, df_input_metadata$Treatment, df_input_metadata$DPE, sep = "_")
df_input_metadata$Strain_Treatment_DPE = factor(df_input_metadata$Strain_Treatment_DPE,
                                              levels = c("AB_Control_Older", "AB_Control_Younger", "AB_Exposed_Older", "AB_Exposed_Younger",
                                                         "5D_Control_Older", "5D_Control_Younger", "5D_Exposed_Older", "5D_Exposed_Younger", 
                                                         "Casper_Control_Older", "Casper_Control_Younger", "Casper_Exposed_Older", "Casper_Exposed_Younger",
                                                         "TU_Control_Older", "TU_Control_Younger", "TU_Exposed_Older", "TU_Exposed_Younger"))
fit_data_all_genus_sta = Maaslin2(input_data     = df_input_data_genus, 
                    input_metadata = df_input_metadata, 
                    min_prevalence = 0,
                    normalization  = "TSS",
                    output         = "fish_output_all_genus_sta", 
                    fixed_effects  = c("Strain_Treatment_DPE"))
View(fit_data_all_genus_sta$results)
```