---
title: "Statistical Analysis"
author: "Emilee Lance"
date: "2024-07-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load Libraries
```{r}
library(phyloseq)
library(vegan)
library(ggplot2)
library(dplyr)
library(MASS)
library(microViz)
library(multcomp)
library(emmeans)
library(FSA)
library(tibble)
library(bestNormalize)
library(kableExtra)
library(stringr)
```
# Create Functions
```{r}
get_formula <- function(model) {
  
  broom::tidy(model)[, 1:2] %>%
    mutate(sign = ifelse(sign(estimate) == 1, ' + ', ' - ')) %>% #coeff signs
    mutate_if(is.numeric, ~ abs(round(., 2))) %>% #for improving formatting
    mutate(a = ifelse(term == '(Intercept)', paste0('y ~ ', estimate), paste0(sign, estimate, ' * ', term))) %>%
    summarise(formula = paste(a, collapse = '')) %>%
    as.character
  
}
```
# Load the Phyloseq object
The phyloseq object was created by following the DADA2 workflow. Documentation can be found here: ./Code/Zebrafish DADA2.Rmd. Phyloseq object saved here: ./Data/ps_object.rds. For this workflow we will be looking only at the Control samples for all 4 strains of zebrafish (5D, AB, Casper & TU).
```{r phyloseq object, tidy = TRUE}
ps <- readRDS("C:/Users/emile/OneDrive - Oregon State University/Sharpton Lab/ZF_ISF__Emiliee_WI2024/Data/ps_object.rds")

ps_exposed <- ps_filter(ps, Treatment == "Exposed")
```
#Extract Sample Data
# Extract Relevant Data
```{r data extraction, tidy = TRUE}
#extract relevant data
#calculate alpha diversity metrics
sam_data <- as.data.frame(ps %>%
 ps_calc_diversity(rank = "Genus", index = "shannon") %>%
 ps_calc_richness(rank = "Genus", index = "observed") %>%
 samdat_tbl())

sam_data <- column_to_rownames(sam_data, var = colnames(sam_data)[1])
```


# Extract Relevant Data and Perform Transformations
```{r data extraction, tidy = TRUE}
#convert Strain to factor
sam_data$Strain <- as.factor(sam_data$Strain)
#add binomial pathology results
sam_data$Pathology.Results.b <- ifelse(sam_data$Pathology.Results == "positive", 1, 0)
#normality for weight
#bestNormalize(sam_data$Weight_mg) #no normalization
#normality for shannon diversity
#test for normality
shapiro_test <- shapiro.test(sam_data$shannon_Genus)
alpha <- 0.05
# Print the results of the Shapiro-Wilk test
print(shapiro_test)
if (shapiro_test$p.value > alpha) {
        print("Sample looks Gaussian/Normally Distributed (fail to reject H0)")
    } else {
        print("Sample does not look Gaussian/Normally Distributed (reject H0)")
    }
bestNormalize(sam_data$shannon_Genus) 
#perform arcsinh transformation
sam_data$arcsinh_shannon_Genus <- asinh(sam_data$shannon_Genus)

#normality for observed counts
shapiro_test <- shapiro.test(sam_data$observed_Genus)
alpha <- 0.05
# Print the results of the Shapiro-Wilk test
print(shapiro_test)
if (shapiro_test$p.value > alpha) {
        print("Sample looks Gaussian/Normally Distributed (fail to reject H0)")
    } else {
        print("Sample does not look Gaussian/Normally Distributed (reject H0)")
    }
# determine best normalization method
bestNormalize(sam_data$observed_Genus) #square root
#perform transformation
sam_data$sqrt_observed_Genus <- sqrt(sam_data$observed_Genus)
#subset
control_data <- subset(sam_data, Treatment == "Control")
exposed_data <- subset(sam_data, Treatment == "Exposed")
```



# Model Selection
## Physiology
Variables to consider: Strain, Sex, and Pathology.Results.
```{r}
#create full model for selection
full_phys_lm <- lm(Weight_mg ~ Strain + Sex + Pathology.Results +Strain*Sex +Strain*Pathology.Results +Sex*Pathology.Results, data = sam_data)
#perform model selection
step(full_phys_lm, direction = "both")
#create new model based on selection results
best_phys_lm <- lm(formula = Weight_mg ~ Strain + Sex + Pathology.Results +Strain*Sex, data = sam_data)
#run analysis of variance
anova(best_phys_lm)
#summary statistics
summary(best_phys_lm)
```
## Alpha Diversity
Variables to consider: Strain, Treatment, DPE, Sex, Weight_mg, Pathology.Result, and Total.Worm.Count.
```{r}
#create shannon diversity full model for selection
full_lm <- lm(arcsinh_shannon_Genus ~ Strain + Treatment + DPE + Sex + Weight_mg + Pathology.Results + Total.Worm.Count + Strain*Treatment + Strain*DPE + Treatment* DPE + Strain*Treatment*DPE, data = sam_data)
#perform model selection
step(full_lm, direction = "both", scope = list(lower = arcsinh_shannon_Genus ~ Strain + Treatment, upper = full_lm))
#create new model based on selection results
shannon_lm <- lm(formula = shannon_Genus ~ Strain + Treatment + DPE + Weight_mg + Treatment:DPE, data = sam_data)
#run analysis of variance
anova(shannon_lm)
#summary statistics
summary(shannon_lm)

# Simple GLM for Strain in each treatment group
shannon_control_lm <- lm(formula = shannon_Genus ~ Strain, data = control_data)
shannon_exposed_lm <- lm(formula = shannon_Genus ~ Strain, data = exposed_data)

#create observed richness full model for selection
full_lm <- lm(sqrt_observed_Genus ~ Strain + Treatment + DPE + Sex + Weight_mg + Pathology.Results + Total.Worm.Count + Strain*Treatment + Strain*DPE + Treatment* DPE + Strain*Treatment*DPE, data = sam_data)
#perform model selection
step(full_lm, direction = "both", scope = list(lower = sqrt_observed_Genus ~ Strain + Treatment, upper = full_lm))
#create new model based on selection results
observed_lm <- lm(formula = sqrt_observed_Genus ~ Strain + Treatment + DPE + 
    Weight_mg + Total.Worm.Count + Strain:Treatment + Strain:DPE + 
    Treatment:DPE, data = sam_data)
#run analysis of variance
anova(observed_lm)
#summary statistics
summary(observed_lm)


# Simple GLM for Strain in each treatment group
observed_control_lm <- lm(formula = observed_Genus ~ Strain, data = control_data)
observed_exposed_lm <- lm(formula = observed_Genus ~ Strain, data = exposed_data)
```

## Beta Diversity
Variables to consider: Strain, Treatment, DPE, Sex, Weight_mg, Pathology.Result, and Total.Worm.Count.
```{r}
#extract the OTU table
otu_table <- otu_table(ps)
#convert OTU table to matrix
otu_matrix <- as(otu_table, "matrix")
#calculate Bray-Curtis dissimilarity matrix
dist_matrix <- vegdist(otu_matrix, method = "bray")
#create full model for selection
full_cap <- capscale(dist_matrix ~ Strain + Treatment + DPE + Sex + Weight_mg + Pathology.Results + Total.Worm.Count + Strain*Treatment + Strain*DPE + Treatment* DPE + Strain*Treatment*DPE, data = sam_data)
#perform stepwise selection
ordistep(full_cap, direction = "both")
#create new model based on model selection results
best_cap <- capscale(dist_matrix ~ Strain + Treatment + DPE + Total.Worm.Count + Strain:Treatment + Strain:DPE +
Treatment:DPE, data = sam_data)
#run adonis2
adonis2(dist_matrix ~ Strain + Treatment + DPE + Total.Worm.Count + Strain:Treatment + Strain:DPE + Treatment:DPE, data = sam_data)
#proportion of variance explained
best_cap$CA$eig / sum(best_cap$CA$eig)

```
# Physiology
```{r}
#Weight by treatment
wilcox.test(Weight_mg ~ Treatment, data =sam_data)
#Weight by pathology results
wilcox.test(Weight_mg ~Pathology.Results, data =sam_data)

#Weight for control
kruskal.test(Weight_mg ~ Strain, data = control_data)
control_dunn_result <- dunnTest(Weight_mg ~ Strain, data = control_data, method = "bonferroni")
print(control_dunn_result)

#Weight for exposed
kruskal.test(Weight_mg ~ Strain, data = exposed_data)
exposed_dunn_result <- dunnTest(Weight_mg ~ Strain, data = exposed_data, method = "bonferroni")
print(exposed_dunn_result)

#Weight by sex
wilcox.test(Weight_mg ~ Sex, data =sam_data)
#Weight by dpe
wilcox.test(Weight_mg ~ DPE, data =sam_data)

#Weight by dpe in exposed
wilcox.test(Weight_mg ~ DPE, data =exposed_data)
#Weight by dpe in control
wilcox.test(Weight_mg ~ DPE, data =control_data)
#Models
# Create Summary Table for best model
phys_table <- anova(best_phys_lm) %>%
  kbl(
    caption = "Weight ~ Intercept + Strain + Sex + Pathology.Results +Strain*Sex",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

anova(lm(Weight_mg ~ Strain + Pathology.Results + Strain*Pathology.Results, data = sam_data))

anova(lm(Weight_mg ~ Strain + Treatment + Strain*Treatment, data = sam_data))

anova(lm(Weight_mg ~ Strain + Pathology.Results + Sex, data = sam_data))


lm(Weight_mg ~ Strain + Pathology.Results + Strain*Pathology.Results, data = sam_data) %>%
  aov() %>%
  TukeyHSD()

#Tukey Test for Best Physiology Model
phys_tukey <- best_phys_lm %>%
  aov() %>%
  TukeyHSD()

# Extract and combine all comparisons
phys_tukey_df <- do.call(rbind, lapply(names(phys_tukey), function(term) {
df <- as.data.frame(phys_tukey[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))

phys_strain_pval_df <- phys_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$shannon_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

phys_strain_tukey_table <- phys_strain_pval_df %>%
  kbl(
    caption = "Post-Hoc Tukey Test for Zebrafish Weight by Strain",
    digits = 3,
    align = "c",
    col.names = c(" ", "lwr", "upr", "diff", "p.adj", "group1", "group2", "term", "y.position", "p.label")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

#Pathology
#infection prevalence
prevalence_results <- sam_data %>%
  filter(Treatment == "Exposed") %>%
  group_by(Strain) %>%
  summarise(
    Total_Exposed = n(),
    Positive_Count = sum(Pathology.Results == "positive"),
    Percent_Positive = (Positive_Count / Total_Exposed) * 100
  )

View(prevalence_results)

prevalence_table <- prevalence_results %>%
  kbl(
    caption = "Infection Prevalence by Strain",
    digits = 2,
    align = "c",
    col.names = c("Strain", "Total Exposed", "Positive Count", "Positive Percent")) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:4, width = "100px")

#by strain
sam_data$Strain <- relevel(sam_data$Strain, ref = "AB")
pathology_model <- lm(Pathology.Results.b ~ Strain, data = exposed_data, family = binomial())
summary(pathology_model)
anova(pathology_model)
posthoc <- glht(pathology_model, linfct = mcp(Strain = "Tukey"))
summary(posthoc)
#now by sex
pathology_model <- lm(Pathology.Results.b ~ Sex, data = exposed_data, family = binomial())
summary(pathology_model)
anova(pathology_model)
# and both
pathology_model <- lm(Pathology.Results.b ~ Strain + Sex, data = exposed_data, family = binomial())

# Create Summary Table for pathology model
pathology_table <- anova(pathology_model) %>%
  kbl(
    caption = "Pathology Outcomes ~ Intercept + Strain + Sex",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")


#Worm Count
# with 5D as reference
exposed_data$Strain<- relevel(exposed_data$Strain, ref = "5D")
glm.nb(Total.Worm.Count ~ Strain + DPE, data = exposed_data) %>%
     summary()

# with AB as reference
exposed_data$Strain<- relevel(exposed_data$Strain, ref = "AB")
glm.nb(Total.Worm.Count ~ Strain+ DPE, data = exposed_data) %>%
 summary()

# with Casper as reference
exposed_data$Strain<- relevel(exposed_data$Strain, ref = "Casper")
glm.nb(Total.Worm.Count ~ Strain+ DPE, data = exposed_data) %>%
 summary()

# with TU as reference
exposed_data$Strain<- relevel(exposed_data$Strain, ref = "TU")
glm.nb(Total.Worm.Count ~ Strain+ DPE, data = exposed_data) %>%
 summary()

burden_lm <- glm.nb(Total.Worm.Count ~ Strain+ DPE, data = exposed_data)

#Tukey test
burden_anova <- glm.nb(Total.Worm.Count ~ Strain + DPE, data = exposed_data) %>%
  anova()
glm.nb(Total.Worm.Count ~ Strain + DPE, data = exposed_data) %>%
  summary()
glm.nb(Total.Worm.Count ~ Strain + DPE, data = exposed_data) %>%
  aov() %>%
  TukeyHSD()

dpe_pval <- burden_anova["DPE", "Pr(>Chi)"]




# Extract and combine all comparisons
burden_anova_df <- do.call(rbind, lapply(names(burden_anova), function(term) {
df <- as.data.frame(burden_anova[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))

# Tukey test
burden_tukey <- burden_lm %>%
  aov() %>%
  TukeyHSD()

# Extract and combine all comparisons
burden_tukey_df <- do.call(rbind, lapply(names(burden_tukey), function(term) {
df <- as.data.frame(burden_tukey[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))

burden_strain_pval_df <- burden_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$shannon_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)


burden_strain_pval_df <- burden_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$shannon_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)


burden_anova_dpe_df <- data.frame(
  group1 = "31",  # replace with actual level name
  group2 = "44",  # replace with actual level name
  p.adj = dpe_pval,
  y.position = 16,
  p.label = case_when(
    dpe_pval <= 0.001 ~ "***",
    dpe_pval <= 0.01 ~ "**",
    dpe_pval <= 0.05 ~ "*",
    dpe_pval <= 0.1 ~ ".",
    TRUE ~ "ns"
  )
)

#Create Summary table
burden_table <- anova(burden_lm) %>%
  kbl(
    caption = "Total Worm Count ~ Intercept + Strain + DPE",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

```
# Factor Effects
## Alpha Diversity
```{r}
#ANOVA, shannon entropy
anova(shannon_lm)
#Tukey Post-hoc test without Weight
shannon_tukey <- lm(formula = shannon_Genus ~ Strain + Treatment + DPE + Treatment:DPE, data = sam_data) %>%
  aov() %>%
  TukeyHSD()


# Extract and combine all comparisons
shannon_tukey_df <- do.call(rbind, lapply(names(shannon_tukey), function(term) {
df <- as.data.frame(shannon_tukey[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))


shannon_strain_pval_df <- shannon_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$shannon_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

shannon_strain_pval_df$Metric <- "Shannon Entropy"

shannon_treatment_pval_df <- shannon_tukey_df %>%
 filter(term == "Treatment") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$shannon_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

shannon_treatment_pval_df$Metric <- "Shannon Entropy"



# Create Summary Table
shannon_table <- anova(shannon_lm) %>%
  kbl(
    caption = "Shannon Entropy ~ Intercept + Strain + Treatment + DPE + Weight_mg + Treatment:DPE",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

#for Control
anova(shannon_control_lm)

shannon_control_tukey <- shannon_control_lm %>%
  aov() %>%
  TukeyHSD()


# Extract and combine all comparisons
shannon_control_tukey_df <- do.call(rbind, lapply(names(shannon_control_tukey), function(term) {
df <- as.data.frame(shannon_control_tukey[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))


shannon_control_strain_pval_df <- shannon_control_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$shannon_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

shannon_control_strain_pval_df$Metric <- "Shannon Entropy"


# Create Summary Table
shannon_control_table <- anova(shannon_control_lm) %>%
  kbl(
    caption = "Shannon Entropy ~ Intercept + Strain",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

#For Exposed Table
anova(shannon_exposed_lm)

shannon_exposed_tukey <- shannon_exposed_lm %>%
  aov() %>%
  TukeyHSD()


# Extract and combine all comparisons
shannon_exposed_tukey_df <- do.call(rbind, lapply(names(shannon_exposed_tukey), function(term) {
df <- as.data.frame(shannon_exposed_tukey[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))


shannon_exposed_strain_pval_df <- shannon_exposed_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$shannon_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

shannon_exposed_strain_pval_df$Metric <- "Shannon Entropy"

# Create Summary Table
shannon_exposed_table <- anova(shannon_exposed_lm) %>%
  kbl(
    caption = "Shannon Entropy ~ Intercept + Strain",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

#ANOVA observed richness
anova(observed_lm)

#Tukey Post-hoc test without Weight or Total.Worm.Count
observed_tukey <- lm(formula = sqrt_observed_Genus ~ Strain + Treatment + DPE +  Strain:Treatment + Strain:DPE + 
    Treatment:DPE, data = sam_data) %>%
  aov() %>%
  TukeyHSD()


# Extract and combine all comparisons
observed_tukey_df <- do.call(rbind, lapply(names(observed_tukey), function(term) {
df <- as.data.frame(observed_tukey[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))


observed_strain_pval_df <- observed_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$observed_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

observed_strain_pval_df$Metric <- "# Observed Genera"

observed_treatment_pval_df <- observed_tukey_df %>%
 filter(term == "Treatment") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$observed_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

observed_treatment_pval_df$Metric <- "# Observed Genera"

alpha_strain_pval_df <- bind_rows(shannon_strain_pval_df, observed_strain_pval_df)

alpha_strain_tukey_table <- alpha_strain_pval_df %>%
  kbl(
    caption = "Post-Hoc Tukey Test for All Samples - Alpha Diversity Metrics",
    digits = 3,
    align = "c",
    col.names = c("diff", "lwr", "upr", "p.adj", "group1", "group2", "term", "y.position", "p.label", "Metric")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1, width = "100px") %>%
column_spec(1:10, width= "65px") %>%
column_spec(11,width = "175px")

alpha_treatment_pval_df <- bind_rows(shannon_treatment_pval_df, observed_treatment_pval_df)

#Create Table
observed_table <- anova(observed_lm) %>%
  kbl(caption = get_formula(observed_lm)) %>%
  kable_classic(full_width = F, html_font = "Cambria")


observed_table <- anova(observed_lm) %>%
  kbl(
    caption = "# Observed Genera ~ Intercept + Strain + Treatment + DPE + Weight_mg + Total.Worm.Count + Strain:Treatment + Strain:DPE + Treatment:DPE",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")


#Tukey Post-hoc test without Weight and worm count
lm(formula = sqrt_observed_Genus ~ Strain + Treatment + DPE + Strain:Treatment + Strain:DPE + Treatment:DPE, data = sam_data) %>%
  aov() %>%
  TukeyHSD()

#for each treatment
anova(observed_control_lm)
anova(observed_exposed_lm)
observed_exposed_lm %>%
  aov() %>%
  TukeyHSD()


#for Control
anova(observed_control_lm)

observed_control_tukey <- observed_control_lm %>%
  aov() %>%
  TukeyHSD()


# Extract and combine all comparisons
observed_control_tukey_df <- do.call(rbind, lapply(names(observed_control_tukey), function(term) {
df <- as.data.frame(observed_control_tukey[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))


observed_control_strain_pval_df <- observed_control_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$observed_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

observed_control_strain_pval_df$Metric <- "# Observed Genera"

# Create Summary Table
observed_control_table <- anova(observed_control_lm) %>%
  kbl(
    caption = "# Observed Genera ~ Intercept + Strain",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

#For Exposed Table
anova(observed_exposed_lm)

observed_exposed_tukey <- observed_exposed_lm %>%
  aov() %>%
  TukeyHSD()


# Extract and combine all comparisons
observed_exposed_tukey_df <- do.call(rbind, lapply(names(observed_exposed_tukey), function(term) {
df <- as.data.frame(observed_exposed_tukey[[term]])
df$comparison <- rownames(df)
df$term <- term
df
}))


observed_exposed_strain_pval_df <- observed_exposed_tukey_df %>%
 filter(term == "Strain") %>%
 rename(p.adj = `p adj`) %>% # Rename the column to avoid issues
 separate(comparison, into = c("group1", "group2"), sep = "-") %>%
mutate(
y.position = max(sam_data$observed_Genus, na.rm = TRUE) + seq(0.1, 0.5, length.out = n()),
 p.label = case_when(
 p.adj <= 0.001 ~ "***",
 p.adj <= 0.01 ~ "**",
 p.adj <= 0.05 ~ "*",
 p.adj <= 0.1 ~ ".",
 TRUE ~ "ns"
 )
)

observed_exposed_strain_pval_df$Metric <- "# Observed Genera"

# Create Summary Table
observed_exposed_table <- anova(observed_exposed_lm) %>%
  kbl(
    caption = "# Observed Genera ~ Intercept + Strain",
    digits = 3,
    align = "c",
    col.names = c("Df", "Sum Sq", "Mean Sq", "F value", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

# Combine both alpha metrics for both control and exposed
alpha_control_pval_df <- bind_rows(shannon_control_strain_pval_df, observed_control_strain_pval_df)


alpha_control_tukey_table <- alpha_control_pval_df %>%
  kbl(
    caption = "Post-Hoc Tukey Test for Control Sample Alpha Diversity Metrics",
    digits = 3,
    align = "c",
    col.names = c("diff", "lwr", "upr", "p.adj", "group1", "group2", "term", "y.position", "p.label", "Metric")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1, width = "100px") %>%
column_spec(1:10, width= "65px") %>%
column_spec(11,width = "175px")


alpha_exposed_pval_df <- bind_rows(shannon_exposed_strain_pval_df, observed_exposed_strain_pval_df)

alpha_exposed_tukey_table <- alpha_exposed_pval_df %>%
  kbl(
    caption = "Post-Hoc Tukey Test for Exposed Sample Alpha Diversity Metrics",
    digits = 3,
    align = "c",
    col.names = c("diff", "lwr", "upr", "p.adj", "group1", "group2", "term", "y.position", "p.label", "Metric")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1, width = "100px") %>%
column_spec(1:10, width= "65px") %>%
column_spec(11,width = "175px")
```
## Beta Diversity
```{r}
#adonis
adonis.result <- adonis2(dist_matrix ~ Strain + Treatment + DPE + Total.Worm.Count + Strain:Treatment + Strain:DPE + Treatment:DPE, data = sam_data)
adonis.result
adonis.result$aov.tab
#run PERMANOVA based on best model
beta_all <- ps %>%
   tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
   tax_agg(rank = "Genus") %>%
   dist_calc(dist = "bray") %>%
   dist_permanova(variables = c("Strain", "Treatment", "DPE", "Total.Worm.Count", interactions = c("Strain*Treatment", "Strain*DPE", "Treatment*DPE")), n_perms = 999, seed = 123) %>%
   perm_get()

# Extract distance matrix and metadata
dist <- ps %>%
   tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
   tax_agg(rank = "Genus") %>%
   dist_calc(dist = "bray") %>%
   dist_get()
 
 
# Run adonis2 directly
beta_all <- adonis2(
   dist ~ Strain + Treatment + DPE + Total.Worm.Count + Strain:Treatment + Strain:DPE + Treatment:DPE,
   data = sam_data,
   permutations = 999,
   by = "terms"  # or try "none"
)

#summary table
beta_table <- beta_all %>%
  kbl(
    caption = "Bray-Curtis Dissimilarity ~ Strain + Treatment + DPE + Total.Worm.Count + Strain:Treatment + Strain:DPE + Treatment:DPE",
    digits = 3,
    align = "c",
    col.names = c("Factor", "Df", "SumsOfSqs", "MeanSqs", "R2", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")


#homogeneity of dispersion tests
bdisp_all <- ps %>%
    tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
    tax_agg(rank = "Genus") %>%
    dist_calc(dist = "bray") %>%
  dist_bdisp(
    variables = c("Strain", "Treatment", "DPE", "Total.Worm.Count"),
    method = c("centroid", "median")[[1]],
    complete_cases = TRUE,
    verbose = TRUE) %>%
  bdisp_get(
)

#Strain Dispersion
bdisp_all$Strain$model$group.distances #average distance to the centroid
bdisp_all$Strain$anova #anova results
bdisp_all$Strain$tukeyHSD$group #Tukey dispersion

#Treatment Dispersion
bdisp_all$Treatment$model$group.distances #average distance to the centroid
bdisp_all$Treatment$anova #anova results
bdisp_all$Treatment$tukeyHSD$group #Tukey dispersion

#DPE Dispersion
bdisp_all$DPE$model$group.distances #average distance to the centroid
bdisp_all$DPE$anova #anova results
bdisp_all$DPE$tukeyHSD$group #Tukey dispersion

#Total Worm Count Dispersion
bdisp_all$Total.Worm.Count$model$group.distances #average distance to the centroid
bdisp_all$Total.Worm.Count$anova #anova results
bdisp_all$Total.Worm.Count$tukeyHSD$group #Tukey dispersion



#for just control
beta_control <- ps %>%
  ps_filter(Treatment == "Control") %>%
   tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
   tax_agg(rank = "Genus") %>%
   dist_calc(dist = "bray") %>%
   dist_permanova(variables = c("Strain", "DPE"), n_perms = 999, seed = 123) %>%
   perm_get()

#summary table
beta_control_table <- beta_control %>%
  kbl(
    caption = "Bray-Curtis Dissimilarity ~ Strain + DPE",
    digits = 3,
    align = "c",
    col.names = c("Factor", "Df", "SumsOfSqs", "MeanSqs", "R2", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")

bdisp_cont <- ps %>%
  ps_filter(Treatment == "Control") %>%
    tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
    tax_agg(rank = "Genus") %>%
    dist_calc(dist = "bray") %>%
  dist_bdisp(
    variables = c("Strain", "DPE"),
    method = c("centroid", "median")[[1]],
    complete_cases = TRUE,
    verbose = TRUE) %>%
  bdisp_get(
)


#run PERMANOVA based on best model just exposed
beta_exposed <- ps %>%
  ps_filter(Treatment == "Exposed") %>%
   tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
   tax_agg(rank = "Genus") %>%
   dist_calc(dist = "bray") %>%
   dist_permanova(variables = c("Strain", "DPE", "Total.Worm.Count", interactions = c("Strain*DPE")), n_perms = 999, seed = 123) %>%
   perm_get()

# Extract distance matrix and metadata
dist_exposed <- ps %>%
  ps_filter(Treatment == "Exposed") %>%
   tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
   tax_agg(rank = "Genus") %>%
   dist_calc(dist = "bray") %>%
   dist_get()
 
 
# Run adonis2 directly
beta_exposed <- adonis2(
   dist_exposed ~ Strain + DPE + Total.Worm.Count + + Strain:DPE,
   data = exposed_sam_data,
   permutations = 999,
   by = "terms"  # or try "none"
)

#summary table
beta_exposed_table <- beta_exposed %>%
  kbl(
    caption = "Bray-Curtis Dissimilarity ~ Strain + DPE + Total Worm Count + Strain*DPE",
    digits = 3,
    align = "c",
    col.names = c("Factor", "Df", "SumsOfSqs", "MeanSqs", "R2", "Pr(>F)")
  ) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1:5, width = "100px")


#homogeneity of dispersion tests
bdisp_exp <- ps %>%
  ps_filter(Treatment == "Exposed") %>%
    tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
    tax_agg(rank = "Genus") %>%
    dist_calc(dist = "bray") %>%
  dist_bdisp(
    variables = c("Strain", "DPE", "Total.Worm.Count"),
    method = c("centroid", "median")[[1]],
    complete_cases = TRUE,
    verbose = TRUE) %>%
  bdisp_get(
)
```










## Strain Effects
```{r}
#alpha diversity
anova(shannon_lm)
#Tukey Post-hoc test without Weight
lm(formula = shannon_Genus ~ Strain + Treatment + DPE + Treatment:DPE, data = sam_data) %>%
  aov() %>%
  TukeyHSD()
## Alpha diversity differs by strain (ANOVA, P = 0.033), with the alpha diversity of the AB and Casper strains being significantly different (Tukey, P = 0.022).

#beta diversity
adonis2(dist_matrix ~ Strain + Treatment + DPE + Total.Worm.Count + Strain:Treatment + Strain:DPE + Treatment:DPE, data = sam_data)
#run PERMANOVA based on best model
ps %>%
   tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
   tax_agg(rank = "Genus") %>%
   dist_calc(dist = "bray") %>%
   dist_permanova(variables = c("Strain", "Treatment", "DPE", "Total.Worm.Count", interactions = c("Strain*Treatment", "Strain*DPE", "Treatment*DPE")), n_perms = 999, seed = 123) %>%
   perm_get()
#homogeneity of dispersion
bdisp_all <- ps %>%
    tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
    tax_agg(rank = "Genus") %>%
    dist_calc(dist = "bray") %>%
  dist_bdisp(
    variables = c("Strain", "Treatment", "DPE", "Total.Worm.Count"),
    method = c("centroid", "median")[[1]],
    complete_cases = TRUE,
    verbose = TRUE) %>%
  bdisp_get(
)

#average distance to the centroid
bdisp_all$Strain$model$group.distances
#anova results
bdisp_all$Strain$anova
#Differences
bdisp_all$Strain$tukeyHSD$group
## Beta Diversity differs between strains (Adonis, P = 0.001). Tukey test reveals significant differences between dispersion of AB and 5D fish (P = 0.033).
```


## Treatment Effects
```{r}
#alpha diversity
anova(shannon_lm)
#Tukey Post-hoc test
lm(formula = shannon_Genus ~ Strain +Treatment + DPE + Treatment:DPE, data = sam_data) %>%
  aov() %>%
  TukeyHSD()
## 

#beta diversity
adonis2(dist_matrix ~ Strain + Treatment + DPE + Total.Worm.Count + Strain:Treatment + Strain:DPE + Treatment:DPE, data = sam_data)
#run PERMANOVA based on best model
ps %>%
   tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
   tax_agg(rank = "Genus") %>%
   dist_calc(dist = "bray") %>%
   dist_permanova(variables = c("Strain", "Treatment", "DPE", "Total.Worm.Count", interactions = c("Strain*Treatment", "Strain*DPE", "Treatment*DPE")), n_perms = 999, seed = 123) %>%
   perm_get()
#homogeneity of dispersion
bdisp_all <- ps %>%
    tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
    tax_agg(rank = "Genus") %>%
    dist_calc(dist = "bray") %>%
  dist_bdisp(
    variables = c("Strain", "Treatment", "DPE", "Total.Worm.Count"),
    method = c("centroid", "median")[[1]],
    complete_cases = TRUE,
    verbose = TRUE) %>%
  bdisp_get(
)

#average distance to the centroid
bdisp_all$Treatment$model$group.distances
#anova results
bdisp_all$Treatment$anova
#Differences
bdisp_all$Treatment$tukeyHSD$group

```

## Other Effects
```{r}
#alpha diversity
anova(shannon_lm)
## Alpha Diversity differs by DPE (ANOVA, P=0.06544), and the interaction between Treatment and DPE (ANOVA, P=0.07436).
summary(shannon_lm)
## Mean alpha diversity scores differ in their relationship to DPE between the control and exposed groups with fish in the exposed group having higher mean alpha diversity scores at 44 DPE and fish in the control group having higher mean alpha diversity scores at 31 DPE (ANOVA, P=0.0744).

#beta diversity
adonis2(dist_matrix ~ Strain + Treatment + DPE + Total.Worm.Count + Strain:Treatment + Strain:DPE + Treatment:DPE, data = sam_data)
#run PERMANOVA based on best model
ps %>%
   tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
   tax_agg(rank = "Genus") %>%
   dist_calc(dist = "bray") %>%
   dist_permanova(variables = c("Strain", "Treatment", "DPE", "Total.Worm.Count", interactions = c("Strain*Treatment", "Strain*DPE", "Treatment*DPE")), n_perms = 999, seed = 123) %>%
   perm_get()

#Strain:Treatment
#create interaction term for manual assessment of Strain:Treatment interaction dispersions
ps@sam_data$"Strain_Treatment" <- paste(ps@sam_data$Strain, ps@sam_data$Treatment, sep = "_")
#homogeneity of dispersion
bdisp_all <- ps %>%
    tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
    tax_agg(rank = "Genus") %>%
    dist_calc(dist = "bray") %>%
  dist_bdisp(
    variables = c("Strain_Treatment"),
    method = c("centroid", "median")[[1]],
    complete_cases = TRUE,
    verbose = TRUE) %>%
  bdisp_get(
)
#average distance to the centroid
bdisp_all$Strain_Treatment$model$group.distances
#anova results
bdisp_all$Strain_Treatment$anova
#Differences
bdisp_all$Strain_Treatment$tukeyHSD$group

# Create a table
# Convert to data frame
tukey_dispersion_df <- as.data.frame(bdisp_all$Strain_Treatment$tukeyHSD$group)
tukey_dispersion_df$comparison <- rownames(tukey_dispersion_df)

# Identify comparisons where both are Control or both are Exposed
tukey_dispersion_df <- tukey_dispersion_df %>%
    mutate(
        group1 = str_extract(comparison, "^[^-]+"),
        group2 = str_extract(comparison, "(?<=-)[^_]+_[^_]+"),
        both_condition = case_when(
            str_detect(group1, "Control") & str_detect(group2, "Control") ~ "Both Control",
            str_detect(group1, "Exposed") & str_detect(group2, "Exposed") ~ "Both Exposed",
            TRUE ~ NA_character_
        )
    )

tukey_dispersion_df <- tukey_dispersion_df %>%
  rename(p_adj = `p adj`) %>%
  rename(Comparison = comparison)


# Filter and highlight
hod_condition_table <- tukey_dispersion_df %>%
  filter(!is.na(both_condition)) %>%
  dplyr::select(Comparison, diff, lwr, upr, p_adj, both_condition) %>%
  kbl(digits = 3, caption = "Comparisons: Both Control or Both Exposed",row.names = FALSE) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1, width = "200px") %>%
column_spec(2:5, width = "50px") %>%
column_spec(6, width = "125px")




# Extract strain names
tukey_dispersion_df <- tukey_dispersion_df %>%
  mutate(
    strain1 = str_extract(group1, "^[^_]+"),
    strain2 = str_extract(group2, "^[^_]+"),
    same_strain = if_else(strain1 == strain2, TRUE, FALSE)
  )

tukey_dispersion_df <- tukey_dispersion_df %>%
  rename(Strain = strain1)

# Filter and highlight
hod_strain_table <- tukey_dispersion_df %>%
  filter(same_strain) %>%
  dplyr::select(Comparison, diff, lwr, upr, p_adj, Strain) %>%
  kbl(digits = 3, caption = "Comparisons Within the Same Strain",row.names = FALSE) %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE, background = "white") %>%
  kable_styling(position = "center", bootstrap_options = c("striped", "hover", "condensed"), font_size = 14) %>%
column_spec(1, width = "200px") %>%
column_spec(2:5, width = "50px") %>%
column_spec(6, width = "125px")


#Strain:DPE
#create interaction term for manual assessment of Strain:DPE interaction dispersions
ps@sam_data$"Strain_DPE" <- paste(ps@sam_data$Strain, ps@sam_data$DPE, sep = "_")
#homogeneity of dispersion
bdisp_all <- ps %>%
    tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
    tax_agg(rank = "Genus") %>%
    dist_calc(dist = "bray") %>%
  dist_bdisp(
    variables = c("Strain_DPE"),
    method = c("centroid", "median")[[1]],
    complete_cases = TRUE,
    verbose = TRUE) %>%
  bdisp_get(
)
#average distance to the centroid
bdisp_all$Strain_DPE$model$group.distances
#anova results
bdisp_all$Strain_DPE$anova
#Differences
bdisp_all$Strain_DPE$tukeyHSD$group

#Treatment:DPE
#create interaction term for manual assessment of Strain:DPE interaction dispersions
ps@sam_data$"Treatment_DPE" <- paste(ps@sam_data$Treatment, ps@sam_data$DPE, sep = "_")
#homogeneity of dispersion
bdisp_all <- ps %>%
    tax_filter(min_prevalence = 2.5 / 100, verbose = FALSE) %>%
    tax_agg(rank = "Genus") %>%
    dist_calc(dist = "bray") %>%
  dist_bdisp(
    variables = c("Treatment_DPE"),
    method = c("centroid", "median")[[1]],
    complete_cases = TRUE,
    verbose = TRUE) %>%
  bdisp_get(
)
#average distance to the centroid
bdisp_all$Treatment_DPE$model$group.distances
#anova results
bdisp_all$Treatment_DPE$anova
#Differences
bdisp_all$Treatment_DPE$tukeyHSD$group
```

# Worm Burden Modelling
```{r, warning = FALSE}
#alpha diversity
exp_sam_data <- sam_data %>% filter(Treatment == "Exposed")
#total worm count visualization
exp_sam_data %>%
     ggplot(aes(y = DPE, x = Total.Worm.Count)) +
     geom_boxplot(width = 0.7) +
     geom_jitter(height = 0.2, aes(color = Strain)) +
     theme_bw()
#scatterplot of richness vs total worm count
exp_sam_data %>%
     ggplot(aes(y = observed_Genus, x = Total.Worm.Count)) +
     geom_point(aes(color = Strain)) +
     theme_bw()
#negative binomial regressions
glm.nb(Total.Worm.Count ~ DPE, data = exp_sam_data) %>%
 summary()
glm.nb(observed_Genus ~ Total.Worm.Count, data = exp_sam_data) %>%
 summary()
glm.nb(shannon_Genus ~ Total.Worm.Count, data = exp_sam_data) %>%
 summary()

```