---
title: "Zebrafish DADA2"
author: "Emilee Lance"
date: "2024-01-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load Libraries
```{r libraries, echo = FALSE, warning = FALSE}
library(knitr)
library(BiocStyle)
library(ggplot2)
library(gridExtra)
library(dada2)
library(phyloseq)
library(DECIPHER)
library(phangorn)
library(tidyverse)
library(microViz)
library(ggtext)
library(ggraph)
library(DT)
library(dplyr)
library(rcompanion)
```

##Load and format the Fastq Files
These were obtained from the CQLS infrastructure.
Files found here: /nfs2/hts/miseq/2021/211018_M01498_0845_000000000-JYH2G/L1
```{r Fastq files, warning = FALSE}
path <- "C:/Users/emile/Documents/OSU/Sharpton Lab/Winter2024_Project/FastQ Files"
#sort forward and reverse reads
fnFs <- sort(list.files(path, pattern="_R1_001.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_001.fastq.gz", full.names = TRUE))
#extract sample names
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
sample.names <- sapply(strsplit(basename(sample.names), "-"), `[`, 5)
#quality check
plotQualityProfile(fnFs[1:2])
plotQualityProfile(fnRs[1:2])
```

##Filter and Trim the Fastq Files
Forward and reverse files are matched, trimmed based on quality of reads with a maximum error rate of 2, then filtered. Trimmed and filtered files are placed in a new folder.
```{r filter and trim, echo = FALSE, warning = FALSE}
filt_path <- file.path(path, "filtered") 
if(!file_test("-d", filt_path)) dir.create(filt_path)
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(210,190),
                      maxN=0, maxEE=2, truncQ=2, rm.phix=TRUE,
                      compress=TRUE, multithread=FALSE)
head(out)
dim(out)
```

##Evaluate Error Rates
```{r errors}
#Learn error rates
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)

#Subset existing file names
filtFs <- filtFs[file.exists(filtFs)]
filtRs <- filtRs[file.exists(filtRs)]

#Sample Inference Data
dadaFs <- dada(filtFs, err=errF, multithread=TRUE)
dadaRs <- dada(filtRs, err=errR, multithread=TRUE)
dadaFs[[1]]


```

##Merge Pairs and Remove Chimeras
```{r pairs}
#Merge Paired Reads
mergers <- mergePairs(dadaFs, filtFs, dadaRs, filtRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])

#Construct Sequence Table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
table(nchar(getSequences(seqtab)))

#Remove chimeras
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
sum(seqtab.nochim)/sum(seqtab)
```

##Look at Sample Reads Throughout the Pipeline
```{r, warning = FALSE}
#Track samples
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(dadaRs, getN), sapply(mergers, getN), rowSums(seqtab.nochim))
colnames(track) <- c("input", "filtered", "denoisedF", "denoisedR", "merged", "nonchim")
rownames(track) <- sample.names
head(track)
```

##Assign Taxonomy
Using the Silva species (version 138) from Zenodo (https://zenodo.org/records/4587955)
```{r Taxonomy}
taxa <- assignTaxonomy(seqtab.nochim, "C:/Users/emile/Documents/OSU/Sharpton Lab/ZF_ISF__Emiliee_WI2024/Data/silva_nr99_v138.1_train_set.fa.gz", multithread=TRUE)
taxa <- addSpecies(taxa, "C:/Users/emile/Documents/OSU/Sharpton Lab/ZF_ISF__Emiliee_WI2024/Data/silva_species_assignment_v138.1.fa.gz")
taxa.print <- taxa
rownames(taxa.print) <- NULL
head(taxa.print)
```
##Preprocess Metadata
```{r Preprocess Metadata}
#load the metadata
samdf <- read.csv("C:/Users/emile/Documents/OSU/Sharpton Lab/ZF_ISF__Emiliee_WI2024/Data/metadata.csv",header=TRUE)

#Weights
##Check for missing weight values and get the metadata
samdf %>%
  filter(!complete.cases(.))
samdf_w <- samdf[!is.na(samdf$Weight_mg), ]

##visualize the distribution of similar samples
samdf %>%
 filter(Strain == "TU" & Treatment == "Control") %>%
 ggplot(aes(x = Weight_mg)) +
 geom_histogram() +
 theme_bw() 

##get summary statistics
summary_stats <- samdf_w %>%
   filter(Strain == "TU" & Treatment == "Control") %>%
   summarise(
       count = n(),
       mean = mean(Weight_mg, na.rm = TRUE),
       sd = sd(Weight_mg, na.rm = TRUE),
       min = min(Weight_mg, na.rm = TRUE),
       max = max(Weight_mg, na.rm = TRUE),
       median = median(Weight_mg, na.rm = TRUE),
       IQR = IQR(Weight_mg, na.rm = TRUE)
   )
print(summary_stats)

#test normality
# Extract the filtered data for normality test
filtered_data <- samdf_w %>%
  filter(Strain == "TU" & Treatment == "Control")

# Apply the Shapiro-Wilk test to the 'Weight_mg' column
shapiro_test <- shapiro.test(filtered_data$Weight_mg)

# Print the results of the Shapiro-Wilk test
print(shapiro_test)
if (shapiro_test$p.value > alpha) {
        print("Sample looks Gaussian/Normally Distributed (fail to reject H0)")
    } else {
        print("Sample does not look Gaussian/Normally Distributed (reject H0)")
    }

#If Values are normally distributed, take the mean, if not, take the median
print(summary_stats$mean)

replacement <- samdf %>%
  filter(!complete.cases(.)) %>%
  replace(.,"Weight_mg" ,summary_stats$mean)

samdf[is.na(samdf$Weight_mg), ] <- replacement

```
##Create Phyloseq Object
```{r Phyloseq Object, echo = FALSE}
#reformat rownames
row.names(samdf) <- samdf$Sample
seqtab.nochim_row_names <- sub("_F_filt.fastq.gz$", "", rownames(seqtab.nochim))
rownames(seqtab.nochim) <- seqtab.nochim_row_names
#create the phyloseq object
ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
               sample_data(samdf), 
               tax_table(taxa))
View(ps@tax_table)

#Add supersaturation information
supersat <- rownames(sample_data(ps))[sample_data(ps)$Strain %in% c("AB", "5D") & sample_data(ps)$Treatment == "Exposed"]
super_sat_values <- ifelse(rownames(sam_data(ps)) %in% supersat, "Y", "N")
sample_data(ps)$SuperSat <- super_sat_values

#add age information
ps <- ps %>%
  microViz::ps_mutate(
    Age = case_when(
      Age.Days > 160 ~ "Older",
      Age.Days < 160 ~ "Younger"
      ))
View(ps@tax_table)
nrow(ps@tax_table)

#Create new DPE column
ps <- ps %>%
  microViz::ps_mutate(
    DPE = case_when(
      Experiment == "31 dpe" ~ 31,
      Experiment == "44 dpe" ~ 44
      ))
ps@sam_data$DPE <- as.factor(ps@sam_data$DPE)

#filter
ps <- ps %>%
  # Any taxa not found in at least 2 samples are removed
  microViz::tax_filter(min_prevalence = 2, undetected = 0) %>%
  # Remove any unwanted reads
  microViz::tax_select(c("Mitochondria", "Chloroplast", "Eukaryota"), deselect = TRUE) %>%
  microViz::tax_select(c("Bacteria, Phylum"), deselect = TRUE)
nrow(ps@tax_table)

#Add ASVs
ps <-
ps %>% 
  tax_name(
    prefix = "ASV",
    rank = NA,
    pad_number = TRUE,
    sep = "_"
  )

ps <- ps %>%
# Fix names for taxonomic ranks not identified
  microViz::tax_fix(suffix_rank = "current", anon_unique = T, unknown = NA)

View(ps@tax_table)
nrow(ps@tax_table)

# save the object for future use
saveRDS(ps, file = "C:/Users/emile/Documents/OSU/Sharpton Lab/ZF_ISF__Emiliee_WI2024/Data/ps_object.rds")
```

##Store DNA Sequences Elsewhere
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
ps
```

##Visualize Most Abundant Taxa
```{r, warning = FALSE}
#sort in order of abundance
top20 <- names(sort(taxa_sums(ps), decreasing=TRUE))[1:20]
ps.top20 <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
ps.top20 <- prune_taxa(top20, ps.top20)
#Plot by strain and separate by treatment
ps.top20.glom.fam <- tax_glom(ps.top20, taxrank = "Family")
plot_bar(ps.top20.glom.fam, x="Strain", fill="Family") + facet_wrap(~Treatment)
ps.top20.glom.cla <- tax_glom(ps.top20, taxrank = "Class")
plot_bar(ps.top20.glom.cla, x="Strain", fill="Class") + facet_wrap(~Treatment)
ps.top20.glom.cla <- tax_glom(ps.top20, taxrank = "Genus")
plot_bar(ps.top20.glom.cla, x="Strain", fill="Genus") + facet_wrap(~Treatment)
```

By the end of this process we created a phyloseq object and did some basic abundance visualization.
